PMWORKBENCH
=======

**UNDER DEVELOPMENT**

This packages provides template scripts for pharmacometric workflows in R. 

The provided workflows include a combination of .R, .Rmd and .Rwd files. They
integrate Nonmem, PSN and R packages for pharmacometrics.

```{r}
library(pmworkbench)
```

Currently containing: 
```{r}
template_available()
```

## Download templates
Templates, including folder structures, are downloaded using init_project(). The templates are *not* executable files but to be seen as a templates. Modifications and setup are needed. 

Each project types have a read me documentation for what is required to use that particular set, which can be downloaded with get_report_readme(). 

## Demo / tutorial
A demo of how the scripts work together is under development.

## Planned expansions
Plans are to expand the set of templates as more people start using it. Next analysis type will likely be QT analysis (workflow + report) and then PK/PD analyses (probably stratified by therapeutic areas). 
