---
title: "poppk_workflow"
author: "Helena Edlund"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{poppk_workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview
Population pharmacokinetic (PK) analysis is evolving into a routine, but time-consuming, task in drug development. Having workflows, tools, and processes in one place can enable industrialization of these standard activities, ensuring that results are ready in time. 

The set of templates in this workflow aims to facilitate population PK analyses using RStudio as the interface. The workflow integrates multiple R packages to explore and handle PK data. Parameter estimation is assumed to be carried out via PsN and nonmem (from R using NMproject).

The workflow contains a mix of pure R scripts and rmarkdown files to support each step of a generic population PK analysis. The scripts are *templates* and you must tailor them towards your project.

## Generic Population PK analysis workflow

* Exploration of data
    - Data set review: ensure correctness, removal and/or correction of any obvious outliers and/or errors
    - Data analysis: Concentration-time graphics, distributions and summaries of covariates, dose-proportionality assessments 

* Modelling & Simulation
    - Generation of model/simulation scripts
    - Model execution and monitoring
    - Model evaluation / post-processing
    - Simulations

* Summarizing results in report and/or presentations: covered in a separate set of templates

# Getting started

1. Initiate a new project e.g. using the functionality of `tidyproject`. Among other things, this creates a certain folder scaffold, e.g. a sub-folder for scripts. You can also create a new RStudio project or folder to work from.

2. Download the `poppk_workflow`

```{r}
library(pmworkbench)
```

```{r, eval=F}
template_download(path = "Scripts",
                  template_type =  "poppk_workflow") 
```

This will give you a local copy of these scripts and additional folders into your 'Scripts' directory: 

```{r, echo=F}
list_files(path = file.path("..",'inst','poppk_workflow'))

```

# Description of each script

## Setup directory
The 'Setup' folder contains files working as a configuration for the project. The purpose of these script is to setup the your R environment and define project specific variables which can be used across the other scripts. Meaning, if carefully used, you will only have to define things once.

### `setup01_rEnvironment.R`



### `setup02_directories.R`



### `setup03_variables.R`


## Functions directory
This folder is intentionally left emtpy but serves as a placeholder for any selfwritten, project specific fundtions you may want to add. Files in here are 



## Main script directory

### `s00_main.R`

This script should mainly be considered a log book for the analysis and is there for reproducibility purposes. It sources/renders the other scripts in the correct order. Hence, once the appropriate modifications have been done to the template scripts, the entire analysis is ``executable'' from s00_main.R



### `s01_dataset_preparation.Rmd`


### `s02_dataset_review.R`

The intention of this script is pure data review: 
 - Is all data requested included in the delivered file?
 - Is the data sorted correctly?
 - Are the derived covariates calculated correctly?
 - Cross-checking of columns (EVID, DV, MDV, AMT, etc.)

### `s03_summary_excluded_data.R`
Summarizes any *a priori* excluded data (commented out). 
Outputs two csv files commonly requested from regulatory agencies: 
  - Summary of excluded data by dose or study
  - Listing of all rows excluded

### `s04_eda_covariates.R`
Numeric summaries of baseline values
Matrix plots of correlations and distributions
Graphics for variations with time


### `s05_eda_conc_time.R`
Concentrations vs. time graphics and summaries
 - Grouped by study and/or dose
 - Individual plots 

### `s06_dose_dependency.R`

Linear and log-log fits + graphics and summaries