---
title: "Dataset review"
author: "Author"
date: "Date"
output: html_document
---
```{r setup, include=FALSE}
###################################################
# s02_datasetReview.Rmd
# Description: Review of rawdata, identification of a priori outliers etc
# Dependencies: s01_datasetPrep.R
###################################################
library(rprojroot)
knitr::opts_chunk$set(echo=F)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 
# knit in this directory -> everything after this chunk is relative top level dir

# Note: R markdown opens a new R session, your global environment is not available.
```

Load the dataset as prepared by "s01_datasetPrep.R":
```{r, echo=T, message=F}
load(file=file.path("./Scripts","s01.RData"))
# load packages
source(file=file.path("./Scripts","Setup","setup01_projectPackages.R"))
```

This review refers to the full data. That is, data including any rows commented out by "C".

# Structure and dimentions of data
The first 10 rows of the data: 
```{r, echo=T}
head(rawdata, 10)
```

The dimentions of the data:
```{r, echo=T}
dim(rawdata)
```

Included columns: 
```{r, echo=T}
names(rawdata)
```

Are all columns in the dataset included and defined in "setup03_projectVariables.R"?
```{r}
if(all(names(rawdata) %in% allCols)){
  print("Yes")
}else{
  paste("No: ", names(rawdata)[!(names(rawdata) %in% allCols)], "are not included.", collapse="")
}
```

Are any columns included that you were not expecting/including in "setup03_projectVariables.R"?
```{r}
if(all(allCols %in% names(rawdata))){
  print("No")
}else{
 paste("Yes: ", allCols[!(allCols %in% names(rawdata))], "are not included.", collapse="")
}
```

Are there any duplicated rows?
```{r}
if(!unique(duplicated(rawdata))){
  "No" 
}else{
  "Yes"
}
```

# Check of included studies, cohorts, parts, doses and subjects
Which studies are included and are they defined in "setup03_projectVariables.R": 
```{r}
unique(rawdata$OSTUDYID)
unique(unique(rawdata$OSTUDYID) %in% ostudies)
```
 
Same for the numeric versions of the study flag:
```{r}
unique(rawdata$STUDYID)
unique(unique(rawdata$STUDYID) %in% studies)
```

Are the original study names and the numeric flag column in alignment?
```{r} 
rawdata %>% 
  group_by(OSTUDYID) %>% 
  summarize(studyid = unique(STUDYID))
```

Which cohorts are included and are they defined in "setup03_projectVariables.R": 
```{r}
unique(rawdata$COHORT)
unique(unique(rawdata$COHORT) %in% cohorts)
```

```{r, include=F}
# Move out and include if parts are in dataset: 
# Which parts are included and are they defined in "setup03_projectVariables.R": 
# unique(rawdata$PART)
# unique(unique(rawdata$PART) %in% parts)
```

Which dose groups are included and are they defined in "setup03_projectVariables.R": 
```{r}
unique(rawdata$DOSE)  
unique(unique(rawdata$DOSE) %in% doses)
```

Cross-check of studies, cohorts, parts and dosing regimens
```{r} 
# code needs to be modified if several dose/freq levels a cohort
rawdata %>% 
  group_by(STUDYID, COHORT) %>% 
  summarize(ostudyid = unique(OSTUDYID),
            studyid  = unique(STUDYID), 
            dose = unique(DOSE),
            freq = unique(FREQ), 
            regimen = unique(paste(DOSE, FREQ)))
```

```{r, include=F} 
# Uncomment and include if dataset has parts
# rawdata %>% 
#   group_by(PART) %>% 
#   summarize(cohort   = unique(COHORT),
#             ostudyid = unique(OSTUDYID),
#             studyid  = unique(STUDYID),
#             dose     = unique(DOSE), 
#             freq = unique(FREQ), 
#             regimen  = unique(paste(DOSE, FREQ)))
```

Does any subject have any (unexpected) change in dosing freqency?
```{r} 
# Don't forget to exclude cohorts with multiple doses/regimens (if any)
tmp <- rawdata %>% 
  group_by(OSTUDYID, OSID) %>% 
  summarize(FREQ = unique(FREQ)) %>% 
  summarize(duplicates=unique(duplicated(OSID))) 
tmp$duplicates # if true there are subject(s) who change
``` 

Does any subject have any (unexpected) change in dose?
```{r} 
# Don't forget to exclude cohorts with multiple doses/regimens (if any)
tmp <- rawdata %>% 
  group_by(OSTUDYID, OSID) %>% 
  summarize(DOSE = unique(!is.na(DOSE))) %>% 
  summarize(duplicates=unique(duplicated(OSID))) 
tmp$duplicates # if true there are subject(s) who change
```

The total number of subjects included is:
```{r}  
nTotal <- length(unique(rawdata$OSID))
nTotal
```

Cross-check of numeric and original subject ID columns:
```{r, echo=T}
nTotal == length(unique(rawdata$NMSEQSID))

rawdata %>% 
  group_by(OSID) %>% 
  summarize(N = length(unique(NMSEQSID))) %>% 
  summarize(max = max(N), 
            duplicates = unique(duplicated(OSID))) # should be 1 and no dupulicates
```

The total number of subjects in each study:
```{r}  
rawdata %>% 
  group_by(OSTUDYID) %>% 
  summarize(N=length(unique(OSID)))
```

Number of PK occasions and maximum follow up time per subject:
```{r}  
rawdata %>% 
  group_by(OSTUDYID, OSID) %>% 
  summarize(nOcc=ifelse(all(is.na(OCC)), 
                        0, max(unique(OCC[is.na(AMT)]), na.rm=T)),
            maxTAFD=max(TAFD, na.rm=T)) %>% 
  summarize(minOccPerId = min(nOcc), 
            maxOccPerId = max(nOcc), 
            minFollowUpPerId = min(maxTAFD), 
            maxFollowUpPerId = max(maxTAFD))
```

# Is the data sorted correctly?
Are the NMSEQSID defined in an increasing order?
```{r} 
ids <- unique(rawdata$NMSEQSID)
idShift <- c(ids[-1], NA)
idDiff <- na.omit(ids-idShift)
all(idDiff < 0)
```

Are the dates and times increasing for each subject?
```{r}
# # This code needs to be double checked so that it works. (no date and time included in this example set)
# rawdata <- 
#   rawdata %>% 
#   mutate(DATETIME = paste(DATE, TIME, sep=" ")) %>% 
#   mutate(DATETIME = as.Date(DATETIME, format = "%Y-%m-%d %H:%M")) # may need to change format
# 
# rawdata %>% 
#   group_by(NMSEQSID) %>% 
#   mutate(DATETIMEShift = c(DATETIME[-1], NA)) %>% 
#   mutate(DATETIMEDiff = (DATETIME-DATETIMEShift)) %>% 
#   summarize(DateTimeIncr = all(DATETIMEDiff[!is.na(DATETIMEDiff)] < 0)) %>% 
#   summarize(unique(DateTimeIncr))
```

Is the TAFD increasing for each subject?
```{r} 
rawdata %>%
  group_by(NMSEQSID) %>%
  mutate(TAFDShift = c(TAFD[-1], NA)) %>%
  mutate(TAFDDiff = (TAFD-TAFDShift)) %>%
  summarize(TAFDIncr = all(TAFDDiff[!is.na(TAFDDiff)] < 0)) %>% 
  summarize(unique(TAFDIncr))
```

Is the TAPD increasing for each dose event?
```{r} 
# Add a flag identifier for each dosing event (can be checked irrespectively of subject id)
tmp <- rawdata
tmp$AMTFLAG[!is.na(tmp$AMT)] <- c(1:nrow(tmp[!is.na(tmp$AMT),]))
# If there are there any rows before the first dose fill those with NA and do locf for the rest
naIndex <- which(!is.na(tmp$AMT))[1]-1
if(naIndex==0){
  tmp$AMTFLAG <- na.locf(tmp$AMTFLAG)
}else{
  tmp$AMTFLAG <- c(rep(NA, naIndex), na.locf(tmp$AMTFLAG))
}
# summarize similar to above
tmp %>%
  group_by(AMTFLAG) %>%
  mutate(TAPDShift = c(TAPD[-1], NA)) %>%
  mutate(TAPDDiff = (TAPD-TAPDShift)) %>%
  summarize(TAPDIncr = all(TAPDDiff[!is.na(TAPDDiff)] < 0)) %>% 
  summarize(unique(TAPDIncr))
```

# Column summaries (not including covariates)
Summary of numeric columns:
```{r}
summary(rawdata[, numericCols])
```

Summary of numeric columns by study:
```{r}
for(i in ostudies){
  print(paste(i,":"))
  print(summary(rawdata[rawdata$OSTUDYID==i, numericCols]))
}
```

Unique values of categorical columns:
```{r}
for(i in factorCols){
  print(rawdata %>% select(i) %>% unique())
}
```

Unique values of categorical columns by study:
```{r}
for(i in factorCols){
print(rawdata %>% group_by(OSTUDYID) %>%  
        select(c("OSTUDYID",i)) %>% unique())
}
```

# Cross-check of AMT, EVID, DV, MDV, BQL and CMT:
Is EVID==0 when AMT is missing?
```{r}
rawdata %>% 
  summarise(evid = unique(EVID[is.na(AMT)])) == 0
```
Is EVID==1 when AMT is *not* missing?
```{r}
rawdata %>% 
  summarise(evid = unique(EVID[!is.na(AMT)])) == 1
```

Is EVID=0 for observations?
```{r}
rawdata %>% 
  summarize(evid = unique(EVID[!is.na(DV)])) == 0  
```

Is BLQ=0 for non-missing observations?
```{r}
rawdata %>% 
  summarize(blq = unique(BLQ[!is.na(DV)])) == 0   
```

Are any quantified concentrations reported $<$LLOQ?
```{r}
rawdata %>% 
  summarize(blqDV = unique(DV[!is.na(DV)] < LLOQ)) 
```

Are all TAPD=0 dosing events (evid=1)?
```{r}
rawdata %>% 
  summarize(evid = unique(EVID[!is.na(TAPD) & TAPD==0])) == 1   
```

Are the compartments set correctly?
```{r}
rawdata %>% 
  group_by(EVID) %>% 
  summarize(cmt = unique(CMT)) 
```

Cross-check columns for rows with missing DV:
```{r}
rawdata %>% 
  subset(is.na(DV) & is.na(AMT)) %>% 
  summarize(dv = unique(DV), 
            evid = unique(EVID), 
            mdv = unique(MDV), 
            blq = unique(BLQ))
```

Does DV and LNDV align?
1. are the same rows missing? 
```{r}
unique(is.na(rawdata$DV) == is.na(rawdata$LNDV)) 
```

2. Is exp(LNDV) the same as DV? If not, what is the max difference? 
```{r}
unique(signif(rawdata$DV[!is.na(rawdata$DV)], digits=5) == 
         signif(exp(rawdata$LNDV[!is.na(rawdata$DV)]), digits=5))
tmp <- 
  data.frame(dv = signif(rawdata$DV[!is.na(rawdata$DV)], digits=5),
             lndv = signif(exp(rawdata$LNDV[!is.na(rawdata$DV)]), digits=5), 
             dev = (signif(rawdata$DV[!is.na(rawdata$DV)], digits=5) - 
                      signif(exp(rawdata$LNDV[!is.na(rawdata$DV)]), digits=5)))
max(abs(tmp$dev))
```

# Cross-check of C and COMMENTS
Is there any C="C" which does *not* have an exclusion comment?
```{r}
rawdata %>% 
  filter(C=="C") %>% 
  summarise(comment = any(COMMENT %in% c("", " ") | is.na(COMMENT)))
```

# Check of concentration-time data
Comparison of nomial time and time after dose: 

Note: I did not include a NOMTIME in this example so this is not included here. The code below is from acalabrutinib project and needs to be checked so that it works properly.
```{r, echo=T}
# data subset with observations only
rawdataConc <- rawdata %>% 
  subset(EVID==0) %>% 
  subset(!(MDV==1 & is.na(BLQ)))

# # This comparison may have to be split by study if data includes many subjects and different study designs
# 
# rawdataConc %>% 
#   mutate(DIFF = NOMTIME - TAPD)
# summary(rawdataConc$DIFF)
```

Are there any missings for NOMTIME or TAPD?
```{r, echo=T}
# tmp <- rawdataConc %>% 
#   subset(is.na(DIFF))
# tmp
# unique(tmp$COMMENT)
# unique(tmp$C)
```

Check of negative differences (i.e. assessment later than planned)
```{r, echo=T}
# rawdataConc %>% 
#   subset(!is.na(DIFF) & DIFF <= -0.17) %>% #>= 10 min difference
#   select(c("OSID","DATE","TIME",'TAFD','NOMTIME','TAPD',
#            "DIFF",'DV','BLQ','FREQ','COMMENT','MDV','OCC','COHORT')) %>% 
#   group_by(FREQ, NOMTIME, COHORT) %>% 
#   summarize(min = min(DIFF),
#             max = max(DIFF),
#             median=median(DIFF),
#             n = length(DIFF))
```

Check of positive differences (i.e. assessment earlier than planned)
```{r, echo=T}
# rawdataConc %>% 
#   subset(!is.na(DIFF) & DIFF >= 0.17) %>% #>= 10 min difference
#   select(c("OSID","DATE","TIME",'TAFD','NOMTIME','TAPD',
#            "DIFF",'DV','BLQ','FREQ','COMMENT','MDV','OCC','COHORT')) %>% 
#   group_by(FREQ, NOMTIME, COHORT) %>% 
#   summarize(min = min(DIFF),
#             max = max(DIFF),
#             median=median(DIFF),
#             n = length(DIFF))
```

# Covariate summary and cross-check
## Baseline data
```{r}
baselineData <- rawdata %>% 
  subset(!duplicated(OSID))
```
Summary of continous covariates
```{r}
summary(baselineData[, bContCov])
``` 
Unique values of categorical covariates
```{r}
apply(baselineData[, bCatCov], 2, unique)
```

Is there only one value for each subject? (should return TRUE for all subjects)
```{r}
perId <- function(x){
  return(length(unique(x))==1) # each individual should only have one value
}
summary(
  rawdata %>% 
  group_by(OSID) %>% 
  summarize_at(c(bContCov,bCatCov), 
               funs(values=perId))
  )
```

### Cross-check of categories to ensure that the coding is correct:
Does males have higher body weight?
```{r}
baselineData %>% 
  subset(BWT!=-99) %>% 
  group_by(SEXM) %>% 
  summarise(min = min(BWT),
            median = median(BWT),
            max = max(BWT))
```

Is the RENAL impairment correct compared to CRCL and/or EGFR?
```{r, echo=T}
baselineData %>% 
  group_by(BRENAL) %>% 
  subset(BEGFR !=-99) %>% 
  summarize(min = min(BEGFR, na.rm = T),
            median = median(BEGFR, na.rm = T),
            max = max(BEGFR, na.rm = T))
# baselineData %>% 
#   group_by(BRENAL) %>% 
#   subset(BCRCL !=-99) %>% 
#   summarize(min = min(BCRCL, na.rm = T),
#             median = median(BCRCL, na.rm = T),
#             max = max(BCRCL, na.rm = T))

# Compare to:
# Normal: 	  eGFR (mL/min/1.73 m2) ≥90; or CRCL (mL/min) ≥90
# Mild:		    eGFR (mL/min/1.73 m2) 60-89; or CRCL (mL/min) 60-89
# Moderate:	  eGFR (mL/min/1.73 m2) 30-59; or CRCL (mL/min) 30-59
# Severe:		  eGFR (mL/min/1.73 m2) 15-29; or CRCL (mL/min) 15-29
# End Stage: 	eGFR (mL/min/1.73 m2) <15 /on dialysis; or CRCL (mL/min) <15 /on dialysis
```

Is the HEPATIC impairment correct compared to BI/AST (if NCI criteria)?
```{r} 
# baselineData %>% 
#   group_by(BHEPATIC) %>% 
#   subset(BBI !=-99) %>% 
#   summarize(minBI = min(BBI, na.rm = T),
#             medianBI = median(BBI, na.rm = T),
#             maxBI = max(BBI, na.rm = T),
#             minAST    = min(BAST, na.rm = T),
#             medianAST = median(BAST, na.rm = T),
#             maxAST    = max(BAST, na.rm = T))

# NCI criteria:
# Normal:    Bilirubin ≤ ULN, and AST ≤ ULN
# Mild:		   Bilirubin ≤ ULN, and AST > ULN or Bilirubin >1.0*ULN to ≤1.5*ULN, and AST of any value
# Moderate:	 Bilirubin > 1.5*ULN to ≤ 3*ULN, and AST of any value
# Severe:	   Bilirubin >3.0*ULN, and AST of any value
# where the upper limit of normal (ULN) for bilirubin and AST
# are defined within the SDTM of each study.
```

Spot check of derived variables:
```{r, echo=T} 
# Selection of random ids to recalculate for
ids <- sample(unique(baselineData$OSID), 10)
temp <- baselineData[baselineData$OSID %in% ids,]
```

```{r, include=F}
# uncomment and include if BCRCL in dataset
# # Re-calculate CRCL and compare
# BCLCR <- signif(
#   ifelse(temp$SEXM==1, ((140-temp$AGE)*(temp$BWT) / (temp$BCREA*0.8136)), # if male
#          0.85*((140-temp$AGE)*(temp$BWT) / (temp$BCREA*0.8136))), # if female
#   digits=3)
# BCLCR
# signif(temp$BCRCL, digits = 3)
```

Re-calculate BEGFR and compare
```{r}
BEGFR <- vector("numeric", length(ids))
for(i in 1:length(ids)){

  # not black male
  BEGFR[i] <- 175 * temp$BSCR[i]^(-1.154) * temp$AGE[i]^(-0.203)

  # if female
  if(temp$SEXM[i]==0){
    BEGFR[i] <- BEGFR[i] * 0.742
  }
  # if black
  if(temp$RACE[i]==2){
    BEGFR[i] <- BEGFR[i] * 1.212
  }
}
signif(BEGFR, digits = 3)
signif(temp$BEGFR, digits = 3)
```

```{r}
# uncomment and include if BBSA in dataset
# Re-calculate BBSA and compare
# BBSA <- signif(0.007184 * (temp$BWT^0.425) * (temp$BHT^0.725), digits=3)
# BBSA
# signif(temp$BBSA, digits = 3)
```

Re-calculate BBMI and compare
```{r}
BBMI <- signif(temp$BWT / ((temp$BHT/100)^2), digits=3)
BBMI
signif(temp$BBMI, digits = 3)
```
