---
title: "Summary of excluded and commented data"
author: "Author"
date: "Date"
output: html_document
params:
  printResults: FALSE
---
```{r setup, include=FALSE}
###################################################
# s03_summaryOfExcludedData.R
# Description: Summary of data excluded with C="C
# Dependencies: s01_datasetPrep.R
###################################################
library(rprojroot)
knitr::opts_chunk$set(echo=F)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 
# knit in this directory -> everything after this chunk is relative top level dir

# Note: R markdown opens a new R session, your global environment is not available.
```

Load the dataset as prepared by "s01_datasetPrep.R":
```{r, echo=T, message=F}
load(file=file.path("./Scripts","s01.RData"))
# load packages
source(file=file.path("./Scripts","Setup","setup01_projectPackages.R"))
```

```{r}
# The output should be: 
# 1. a summary table of number of excluded data and reasons (split by >/< LLOQ)
# 2. a table listing each excluded record
# Both to be used in the report. 
```

# Summary of excluded subjects, concentrations and doses
Number of subjects excluded and total number in final data:
```{r}
nTotal <- length(unique(rawdata$OSID))
nIncl <- length(unique(data$OSID))
nIncl - nTotal
nIncl
```

Total excluded rows: 
```{r}
excluded <- rawdata[rawdata$C=="C", ]
nrow(excluded)

exclDoses <- excluded[excluded$EVID==1,]
exclConc <- excluded[excluded$EVID==0,]
```
Out of which `r nrow(exclDoses)` records are doses and `r nrow(exclConc)` concentrations.

Is anything else than EVID=1 or 0 that is excluded?
```{r}
! sum(nrow(exclDoses),nrow(exclConc)) ==  nrow(excluded)
```

How much of the excluded concentration data is missing or BLQ?
```{r}
exclConcNoBQL <- exclConc[exclConc$MDV==1,]
paste0(nrow(exclConcNoBQL)," (", round(100*(nrow(exclConcNoBQL) / nrow(exclConc))), "%)")
```

# Review and summary of comments
## Excluded doses
```{r}
unique(exclDoses[,"COMMENT"])
```

## Excluded concentrations
```{r}
unique(exclConc[,"COMMENT"])
```
### Reason for exclusion split by BLQ/non-BLQ:
```{r}
# In this dataset there are not multiple comments in one row/cell. 
# If that is the case, you need to be careful not to count it twice/overwrite the reason.
exclConc$Reason[str_detect(exclConc$COMMENT, "Pre first dose sample") & 
                  exclConc$BLQ==1] <- "Pre first dose sample BLQ"
exclConc$Reason[str_detect(exclConc$COMMENT, "Pre first dose sample") & 
                  exclConc$BLQ==0] <- "Pre first dose sample Non-BLQ"

exclConc$Reason[str_detect(exclConc$COMMENT, "Randomly selected for illustration") & 
                  exclConc$BLQ==1] <- "Random exclusion BLQ"
exclConc$Reason[str_detect(exclConc$COMMENT, "Randomly selected for illustration") & 
                  exclConc$BLQ==0] <- "Random exclusion Non-BLQ"

totalConc <- rawdata %>% 
  filter(EVID==0)

summary <- 
  exclConc %>%
  group_by(Reason) %>% 
  summarize(n = n(), 
            percentOfTotal = round(100 * (n / nrow(totalConc)), digits=1))
summary
```


```{r, include=F}
if(params$printResults){
  # Output in reportable table
  outTab <- summary
  outTab <- outTab %>% 
    mutate(Values = paste0(n, " (", percentOfTotal, "%)")) %>% 
    select(Reason, Values)
  write.csv(outTab,
            file = paste(resEDADir, "omittedConcDataSummary.csv", sep="/"),
            row.names = F)
}
```

### How much of the non-BLQ data are sparse samples: 
```{r}
# Note, sparse data is those without a OCC assignment in this particular dataset
# Hence code might need to be changed depending on current data
sparse <- 
  exclConc %>%
  filter(is.na(OCC)) %>% 
  group_by(Reason) %>% 
  summarize(n = n(), 
            percentOfTotal = round(100 * (n / nrow(totalConc)), digits=1))
sparse
```


# Review of comments not leading to exclusion
```{r}
# Use the dataset 'data' for this since all C=="C" is commented out
unique(data$COMMENT)

data %>%
  filter(! (COMMENT %in% c(" ", "") | is.na(COMMENT))) %>% 
  group_by(EVID, COMMENT) %>% 
  summarize(n = n(), 
            percentOfTotal = round(100 * (n / nrow(data)), digits=2))
```
