---
title: "Exploratory data analysis: Covariates"
author: "Author"
date: "Date"
output: 
  html_document:
    number_sections: yes
    self_contained: no
    toc: yes
    toc_depth: 2
params:
  print_results: FALSE
---

```{r setup, include=F}
###################################################
# s04_eda_cvariates.Rmd
# Description: Exploratory data analysis of covariates for population PK analysis
# Dependencies: s01_dataset_preparation.R / s01.RData
###################################################
library(rprojroot)
knitr::opts_chunk$set(echo=F)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 
# For correct printing of dust tables
options(pixie_interactive=F)
# Note: R markdown opens a new R session, your global environment is not available.
```

This script uses the data.frame "data", loaded from s01.RData. That is, all rows with C=="C" has been excluded for exploratory data analysis.
```{r environmentPrep, echo=T, message=F}
# ------------------------------------------------------------------
#  Prepare environment
# ------------------------------------------------------------------
# load packages
source(file = file.path("Scripts","Setup","setup01_packages.R")) 
# load output from s01_DatasetPrep.R
load(file=file.path("Scripts","s01.RData"))

# White background in plots
theme_set(theme_bw()) # to be replaced with a azTheme
update_geom_defaults("point", list(shape = 1))
```

Are the plots and tables also being written to file?
```{r}
params$print_results
```

# Baseline covariates
## Numeric summaries
### Entire dataset
The markdown/latex notation for tables does not work for html output. Notified the authour of pixidust.
```{r cont_tab, warning=F, message=F}
# 1. Continuous covariates
continuous_covariates <- 
  summarize_cov(baseline_data, base_cont_cov, type="Cont")

## Collaps to reportable table
cont_tab <- 
  continuous_covariates %>% 
  mutate(Values = paste0(format(Mean, trim=T), " (",
                         format(SD, trim=T),"), [",
                         format(Min, trim=T),"-",
                         format(Max, trim=T) ,"]"),
         Missing = paste0(format(NMissing, trim=T), " (" , 
                          format(PercentMissing, trim=T), ")")) %>% 
  select("Characteristic", "Values", "Missing")

# Labels in this call needs to be set for each specific dataset
cont_tab_markdown <- cont_tab %>% 
  mutate(Characteristic = factor(Characteristic,
                                 levels = c("AGE", "BWT", "BHT","BBMI","BEGFR","BSCR"), 
                                 labels = c("Age (yrs)",
                                            "Body weight (kg)", 
                                            "Body height (cm)",
                                            "Body mass index (kg/m^2^)",
                                            "eGFR (mL/min/1.73m^2^)",
                                            "Serum creatinine (mg/dL)")))

dust(cont_tab_markdown,
     caption = "Continuous covariates") %>% 
  sprinkle_colnames(Values = "mean (SD) [range]", Missing = "Missing (N (%))")
```

```{r, message=F, include=F}
# --------- Save to file
if(params$print_results){
  # if printing to file, use latex notations for lables so that it works with the reporting
  cont_tab_latex <- cont_tab %>% 
    mutate(Characteristic = factor(Characteristic,
                                   levels = c("AGE", "BWT", "BHT","BBMI","BEGFR","BSCR"), 
                                   labels = c("Age ($years$)",
                                              "Body weight ($kg$)", 
                                              "Body height ($cm$)",
                                              "Body mass index ($kg/m^{2}$)",
                                              "eGFR ($mL/min/1.73m^{2}$)",
                                              "Serum creatinine ($mg/dL$)")))
  
  write.csv(cont_tab_latex, row.names = F, 
            file = file.path(res_eda_dir, paste0("cont_tab_", delivery_date,".csv")))
}
```

```{r cat_tab, warning=F, message=F}
categorical_covariates <- 
  summarize_cov(baseline_data, base_cat_cov, type="Cat")

## Collaps to reportable table
cat_tab <- 
  categorical_covariates %>% 
  mutate(Values=paste0(N, " (",
                       format(Percent, trim=T), ")")) %>% 
  select("Characteristic", "Category", "Values")

# Labels in this call needs to be set for each specific dataset
# Use latex notation so that it works with the reporting
cat_tab <- cat_tab %>% 
  mutate(Characteristic = factor(Characteristic, 
                                 levels = c("SEXM","RACE","ETHNIC","BRENAL"), 
                                 labels = c("Sex","Race","Ethnicity","Renal impairment")), 
         Characteristic = as.character(Characteristic), 
         Characteristic = ifelse(duplicated(Characteristic), " ", Characteristic))

dust(cat_tab, 
     caption = "Categorical covariates") %>% 
  sprinkle_colnames(Values = "N (%)")
```

```{r, message=F, include=F}
#--------- Save to file
if(params$print_results){
  write.csv(cat_tab, row.names = F, 
            file=file.path(res_eda_dir, paste0("cat_tab_", delivery_date,".csv")))
} 
```

### Stratified by study
```{r cont_tab_study, warning=F, message=F}

# to be updated with dplyr notation/ put into function... 
for(i in levels(baseline_data$OSTUDYID)){
  # Study subset
  dat <- baseline_data %>% filter(OSTUDYID==i)
  # Study summary
  temp <- summarize_cov(dat, base_cont_cov, type="Cont") %>% 
    mutate(GROUP = i)

  if(i == unique(baseline_data$OSTUDYID)[1]){
    continuous_cov_by_study <- temp
  } else {
    continuous_cov_by_study <- rbind(continuous_cov_by_study, temp)
  }
}

## Collaps to reportable table
cont_tab_by_study <-
  continuous_cov_by_study %>% 
  mutate(Values = paste0(format(Mean, trim=T), " (",
                         format(SD, trim=T),"), [",
                         format(Min, trim=T),"-",
                         format(Max, trim=T) ,"]"),
         Missing = paste0(format(NMissing, trim=T), " (" , 
                          format(PercentMissing, trim=T), ")")) %>% 
  select("GROUP", "Characteristic", "Values", "Missing") %>% 
  rename(Study = GROUP)

# Labels in this call needs to be set for each specific dataset
# Use latex notation so that it works with the reporting
continuous_cov_by_study_markdown <- cont_tab_by_study %>% 
  mutate(Characteristic = factor(Characteristic,
                                 levels = c("AGE","BWT","BHT","BBMI","BEGFR","BSCR"),
                                 labels = c("Age (yrs)",
                                            "Body weight (kg)",
                                            "Body height (cm)",
                                            "BMI (kg/m^2^)",
                                            "eGFR (mL/min/1.73m^2^)",
                                            "Serum creatinine (mg/dL)")), 
         Study = as.character(Study), 
         Study = ifelse(duplicated(Study), "", Study))

dust(continuous_cov_by_study_markdown, 
     caption = "Continuous covariates by study") %>% 
  sprinkle_colnames(Values = "mean (SD) [range]", Missing = "Missing (N (%))")
```

```{r, message=F, include=F}
#--------- Save to file
if(params$print_results){
  # Use latex notation so that it works with the reporting
  cont_tab_by_study_latex <- cont_tab_by_study %>% 
    mutate(Characteristic = factor(Characteristic,
                                   levels = c("AGE","BWT","BHT","BBMI","BEGFR","BSCR"),
                                   labels = c("Age ($years$)",
                                              "Body weight ($kg$)",
                                              "Body height ($cm$)",
                                              "BMI ($kg/m^{2}$)",
                                              "eGFR ($mL/min/1.73m^{2}$)",
                                              "Serum creatinine ($mg/dL$)")), 
           Study = as.character(Study), 
           Study = ifelse(duplicated(Study), "", Study))
  
  write.csv(cont_tab_by_study_latex, row.names = F, 
            file = file.path(res_eda_dir, paste0("cont_tab_by_study_",delivery_date,".csv")))
}
```

```{r cat_tab_study, warning=F, message=F}
# to be updated with dplyr notation/ put into function... 
for(i in levels(baseline_data$OSTUDYID)){
  # Study subset
  dat <- baseline_data %>% filter(OSTUDYID==i)

  temp <- summarize_cov(dat, base_cat_cov, type="Cat") %>% 
    mutate(GROUP = i)

  if(i == unique(baseline_data$OSTUDYID)[1]){
    categorical_cov_by_study <- temp
  } else {
    categorical_cov_by_study <- rbind(categorical_cov_by_study, temp)
  }
}

## Collaps to reportable table
cat_tab_by_study <-
  categorical_cov_by_study %>% 
  mutate(Values=paste0(N, " (",
                       format(Percent, trim=T), ")")) %>% 
  select("GROUP","Characteristic", "Category", "Values") %>% 
  rename(Study = GROUP)

# Labels in this call needs to be set for each specific dataset
# Use latex notation sot that it works with the reporting
cat_tab_by_study <- cat_tab_by_study %>% 
  mutate(Characteristic = factor(Characteristic,
         levels = c("SEXM","RACE","ETHNIC","BRENAL"),
         labels = c("Sex","Race","Ethnicity","Renal impairment")), 
         Characteristic = as.character(Characteristic), 
         Characteristic = ifelse(duplicated(paste(Study, Characteristic)), "", Characteristic), 
         Study = as.character(Study), 
         Study = ifelse(duplicated(Study), "", Study))

dust(cat_tab_by_study, 
     caption= "Categorical covariates by study") %>% 
  sprinkle_colnames(Values = "N (%)")
```

```{r, message=F, include=F}
#--------- Save to file
if(params$print_results){
  write.csv(cat_tab_by_study, 
            file = file.path(res_eda_dir, paste0("cat_tab_by_study_",delivery_date,".csv")),
            row.names = F)
}
```

### Stratified by dose
```{r cont_tab_dose, warning=F, message=F}
# to be updated with dplyr notation/ put into function... 
for(i in levels(baseline_data$DOSE)){
  # Dose group subset
  dat <- baseline_data %>% filter(DOSE == i)
  # Study summary
  temp <- summarize_cov(dat, base_cont_cov, type="Cont") %>% 
    mutate(GROUP = i)

  if(i == unique(baseline_data$DOSE)[1]){
    continuous_cov_by_dose <- temp
  } else {
    continuous_cov_by_dose <- rbind(continuous_cov_by_dose, temp)
  }
}

## Collaps to reportable table
cont_tab_by_dose <- continuous_cov_by_dose %>% 
  mutate(Values = paste0(format(Mean, trim=T), " (",
                         format(SD, trim=T),"), [",
                         format(Min, trim=T),"-",
                         format(Max, trim=T) ,"]"),
         Missing = paste0(format(NMissing, trim=T), " (" , 
                          format(PercentMissing, trim=T), ")")) %>% 
  select("GROUP", "Characteristic", "Values", "Missing") %>% 
  rename(Dose = GROUP) %>% 
  mutate(Dose = as.numeric(Dose)) %>% 
  arrange(Dose)

# Labels in this call needs to be set for each specific dataset
# Use latex notation so that it works with the reporting
cont_tab_by_dose_markdown <- cont_tab_by_dose %>% 
  mutate(Characteristic = factor(Characteristic,
                                 levels = c("AGE","BWT","BHT","BBMI","BEGFR","BSCR"),
                                 labels = c("Age (yrs)",
                                            "Body weight (kg)",
                                            "Body height (cm)",
                                            "BMI (kg/m^2^)",
                                            "eGFR (mL/min/1.73m^2^)",
                                            "Serum creatinine (mg/dL)")), 
         Dose = as.character(Dose), 
         Dose = ifelse(duplicated(Dose),"", Dose))

dust(cont_tab_by_dose_markdown, 
     caption = "Continuous covariates by dose") %>% 
  sprinkle_colnames(Values = "mean (SD) [range]", Missing = "Missing (N (%))")
```

```{r, message=F, include=F}
#--------- Save to file
if(params$print_results){
  # Use latex notation so that it works with the reporting
  cont_tab_by_dose_latex <- cont_tab_by_dose %>% 
    mutate(Characteristic = factor(Characteristic,
                                   levels = c("AGE","BWT","BHT","BBMI","BEGFR","BSCR"),
                                   labels = c("Age ($years$)",
                                              "Body weight ($kg$)",
                                              "Body height ($cm$)",
                                              "BMI ($kg/m^{2}$)",
                                              "eGFR ($mL/min/1.73m^{2}$)",
                                              "Serum creatinine ($mg/dL$)")), 
           Dose = as.character(Dose), 
           Dose = ifelse(duplicated(Dose),"", Dose))
  
  write.csv(cont_tab_by_dose_latex, row.names = F, 
            file = file.path(res_eda_dir, paste0("cont_tab_by_dose_",delivery_date,".csv")))
}
```

```{r cat_tab_dose, warning=F, message=F}
# to be updated with dplyr notation/ put into function... 
for(i in levels(baseline_data$DOSE)){
  # Study subset
  dat <- baseline_data %>% filter(DOSE == i)

  temp <- summarize_cov(dat, base_cat_cov, type="Cat") %>% 
    mutate(GROUP = i)

  if(i == unique(baseline_data$DOSE)[1]){
    categorical_tab_by_dose <- temp
  } else {
    categorical_tab_by_dose <- rbind(categorical_tab_by_dose, temp)
  }
}

## Collaps to reportable table
cat_cov_by_dose <-
  categorical_tab_by_dose %>% 
  mutate(Values=paste0(N, " (",
                       format(Percent, trim=T), ")")) %>% 
  select("GROUP","Characteristic", "Category", "Values") %>% 
  rename(Dose = GROUP) %>% 
  mutate(Dose = as.numeric(Dose)) %>% 
  arrange(Dose)

# Labels in this call needs to be set for each specific dataset
# Use latex notation sot that it works with the reporting
cat_cov_by_dose <- cat_cov_by_dose %>% 
  mutate(Characteristic = factor(Characteristic,
                                 levels = c("SEXM","RACE","ETHNIC","BRENAL"),
                                 labels = c("Sex","Race","Ethnicity","Renal impairment")), 
         Characteristic = as.character(Characteristic), 
         Characteristic = ifelse(duplicated(paste(Dose, Characteristic)), "", Characteristic), 
         Dose = as.character(Dose), 
         Dose = ifelse(duplicated(Dose), "", Dose))

dust(cat_cov_by_dose, 
     caption= "Categorical covariates by dose") %>% 
  sprinkle_colnames(Values = "N (%)")
```

```{r, message=F, include=F}
#--------- Save to file
if(params$print_results){
  write.csv(cat_cov_by_dose, 
            file = file.path(res_eda_dir,
                             paste0("base_cat_cov_by_dose_", delivery_date,".csv")),
            row.names = F)
}
```


# Plots of distributions and correlations
## Entire dataset
### Continuous
The diagonal graphs show histograms of each covariate. The lower off-diagonal graphs are scatter plots of observations (black open circles) with linear regression (black line) and its 95%
confidence interval (grey shaded area). The lower off-diagonal graphs show the Pearson’s correlation coefficient. The graphs are displayed in red if the Pearson’s correlation coefficient is > 0.4.
```{r cont_cont_plot, fig.width=8, fig.height=8, message=F, warning=F}
# You may need to set up different lists if you have many covariates
# or lumping of groups based on too few individuals etc.
cont_vs_cont <-
  ggpairs(baseline_data, columns = base_cont_cov,
          diag = list(continuous = 'barDiag'),
          upper = list(continuous = wrap(ally_cor, 
                                         cor_method = "pearson", 
                                         cor_threshold = 0.4)),
          lower = list(continuous = wrap(ally_scatter_lm_cor, 
                                         cor_method = "pearson", 
                                         cor_threshold = 0.4)))
print(cont_vs_cont)

if(params$print_results){
  ggsave(file=file.path(res_eda_dir, 
                    paste0("continuous_covariates_", delivery_date,".pdf")), 
         plot=cont_vs_cont, 
         height=8, width=8, units = 'in', 
         device=cairo_pdf)
}

```

### Categorical
The diagonal graphs show bar charts of each covariate. The off-diagonal graphs show the correlation between covariate categories: the black point is a visual reference point, and the numbers are percentage of subjects of a covariate split by the groups of the other covariate. For example, the bottom left graph show that within the group with normal renal function, 19% are female and 81% are male (numbers aligned left of the reference point). Similarly, within the group of females, 29% have normal, 36% have mild, 21% have moderate renal impairment and 14% is missing information (numbers aligned above the reference point). NA refers to not available, i.e., missing. 
```{r cat_cat_plot, fig.width=9, fig.height=9, message=F}
cat_vs_cat <-
  ggpairs(baseline_data, columns = base_cat_cov,
          diag = list(discrete = ally_bar),
          upper = list(discrete = "blank"),
          lower = list(discrete = wrap(ally_count, counts="rel"))) 
# if "abs" is used instead of rel you will get the absolute number of patients in each group
print(cat_vs_cat)

if(params$print_results){
  ggsave(file=file.path(res_eda_dir, paste0("categorical_covariates_",delivery_date,".pdf")), 
         plot=cat_vs_cat, 
         height=8, width=8, units = 'in', 
         device=cairo_pdf)
}
```


### Categorical versus continuous
The black line within the box shows the median and the box's upper and lower edges show the inter quartile range (IQR). Whiskers extend to the highest value that is within 1.5*IQR. Data beyond the end of the whiskers are outliers and plotted as points. NA refers to not available, i.e., missing.

```{r cont_cat_plot, fig.width=8, fig.height=10, message=F, warning=F}
cont_vs_cat <-
  ggduo(baseline_data, base_cat_cov, base_cont_cov,
        types = list(comboVertical = ally_box))
print(cont_vs_cat)

if(params$print_results){
  ggsave(file=file.path(res_eda_dir, paste0("cont_vs_cat_covariates_",delivery_date,".pdf")), 
         plot=cont_vs_cat, 
         height=10, width=8, units = 'in', 
         device=cairo_pdf)
}
```

## Distributions stratified by study
```{r cov_study, fig.width=2.5, fig.height=5, message=F, warning=F, include=F}
cont_by_study <-
  ggduo(baseline_data, "OSTUDYID", base_cont_cov,
        types = list(comboVertical = ally_box))
print(cont_by_study)
cat_by_study <-
  ggduo(baseline_data, "OSTUDYID", base_cat_cov,
        types = list(discrete = wrap(ally_count, numbers="rel")))
print(cat_by_study)

if(params$print_results){
  # needs to be pdf if several figures in the same file
  pdf(file=file.path(res_eda_dir, paste0("covariates_by_study_",delivery_date,".pdf")),
      height=5, width=2.5)
  print(cont_by_study)
  print(cat_by_study)
  dev.off()
}
```

## Distributions by dose group/regimen
The black line within the box shows the median and the box's upper and lower edges show the inter quartile range (IQR). Whiskers extend to the highest value that is within 1.5*IQR. Data beyond the end of the whiskers are outliers and plotted as points.

```{r cont_dose, fig.width=3.5, fig.height=6, message=F, warning=F}
cont_by_dose <-
  ggduo(baseline_data, "DOSE", base_cont_cov,
        types = list(comboVertical = ally_box))
print(cont_by_dose)
```

The diagonal graphs show bar charts of each covariate. The off-diagonal graphs show the correlation between covariate categories: the black point is a visual reference point, and the numbers are percentage of subjects of a variable split by the groups of the other variable. NA refers to not available, i.e., missing. See also example text for categorical covariate correlation above.

```{r cat_dose, fig.width=4.5, fig.height=6, message=F, warning=F}
cat_by_dose <-
  ggduo(baseline_data, "DOSE", base_cat_cov,
        types = list(discrete = wrap(ally_count, counts="rel")))
print(cat_by_dose)
```



```{r, include=F}
# write to file
if(params$print_results){
  pdf(file=file.path(res_eda_dir, paste0("covariates_by_dose_", delivery_date,".pdf")),
      height=5, width=2.5)
  print(cont_by_dose)
  print(cat_by_dose)
  dev.off()
}
```

# Time-varying covariates versus time
## Stratified by study
### Continuous
The lines connect data from one individual. Ticks indicate all individual records of the covariate in the dataset. The blue line with shaded area is a loess smooth and its 95% confidence interval indicating any overall trends in changes of covariate values over time after first dose. 

```{r time_var_cont_covs, fig.height=6.5, fig.width=5.5, message=F, warning=F}
# (these covs do actually not change with time but included them as an example)
# If -99/other flag for missing variables is included in the dataset 
# I recommend replacing them with NA for this plot.

# Shortest and longest TAFD to set the same x-axes in plot
tafd_range <- conc_data %>% 
  summarize(min = min(TAFD), 
            max = max(TAFD))

# 1. Continuous cov. vs time by study
p1 <- list()
for(i in cols_cont_cov){
  # Create list with each covariate vs time
  p1[[i]] <-
    ggplot(conc_data, aes_string(x="TAFD", y=i)) +
    geom_point(aes_string(group="OSID"), size = 1.5, shape=124) +
    geom_line(aes_string(group="OSID")) +
    geom_smooth(aes_string(x="TAFD", y=i),
                inherit.aes = F, method="loess") +
    facet_wrap(~OSTUDYID, labeller = "label_both", nrow=2) +
    guides(group="none") +
    coord_cartesian(xlim = c(tafd_range$min, tafd_range$max)) +
    labs(y=i, x = labs_TAFD)
}
invisible(lapply(p1, grob_draw))

if(params$print_results){
  pdf(file=file.path(res_eda_dir, paste0("time_varying_cont_cov_TAFD_", delivery_date,".pdf")),
      height=8.5, width=11)
  invisible(lapply(p1, grob_draw))
  dev.off()
}
```

### Categorical
The lines are a step function connecting data from one individual. Ticks indicate all individual records of the covariate in the dataset. 

```{r time_var_cat_covs, fig.height=6.5, fig.width=5.5, message=F, warning=F}
p1 <- list()
for(i in cols_cat_cov){
  # Create list with each covariate vs time
  p1[[i]] <-
    ggplot(conc_data, aes_string(x="TAFD", y=i, group="OSID")) +
    geom_point(size = 1.5, shape=124) +
    geom_step() +
    facet_wrap(~OSTUDYID, labeller = "label_both", nrow=2) +
    guides(group="none") +
    coord_cartesian(xlim = c(tafd_range$min, tafd_range$max)) + 
    labs(y=i, x = labs_TAFD)
}
invisible(lapply(p1, grob_draw))

if(params$print_results){
  pdf(file=file.path(res_eda_dir,
                 paste0("time_varying_cat_cov_TAFD_", delivery_date,".pdf")),
      height=8.5, width=11)
  invisible(lapply(p1, grob_draw))
  dev.off()
}
```



## Stratified by subject
### Continuous
Ticks indicate all individual records of the covariate in the dataset. 

```{r ind_time_var_cont_cov, fig.height=5.5, fig.width=6.5, warning=F, message=F}
time_var_cont_cov_plots <-  list()

for(i in cols_cont_cov){

  cov_range <- data.frame(
    min = min(conc_data[,i], na.rm=T), 
    max = max(conc_data[,i], na.rm=T))

  # list of individual plots for each covariate
  p1 <- vector("list", length(conc_data_id_splits))
  for(j in 1:length(conc_data_id_splits)){
    p1[[j]] <-
      ggplot(conc_data_id_splits[[j]], aes_string(x="TAFD", y=i)) +
      geom_point(size = 1.5, shape=124) +
      geom_line() +
      facet_wrap(~OSID+OSTUDYID, labeller = "label_both", nrow=3, ncol=4) +
      coord_cartesian(xlim = c(tafd_range$min, tafd_range$max),
                      ylim =c(cov_range$min, cov_range$max)) +
      labs(y=i, x = labs_TAFD)
  }
  # join lists for all covariates
  time_var_cont_cov_plots <- c(time_var_cont_cov_plots, p1)
}

# takes some time to print if many individuals and many covariates...
invisible(lapply(time_var_cont_cov_plots, grob_draw))

if(params$print_results){
  pdf(file=file.path(res_eda_dir,
                 paste0("individual_time_varying_cont_cov_TAFD_", delivery_date,".pdf")),
      height=8.5, width=11)
  invisible(lapply(time_var_cont_cov_plots, grob_draw))
  dev.off()
}
```

## Categorical
The lines are a step function connecting data from one individual. Ticks indicate all individual records of the covariate in the dataset. 

```{r ind_time_var_cat_cov, fig.height=5.5, fig.width=7, warning=F, message=F}
time_var_cat_cov_plots <- list()

for(i in cols_cat_cov){
  # list of individual plots for each covariate
  p1 <- vector("list", length(conc_data_id_splits))
  for(j in 1:length(conc_data_id_splits)){
    p1[[j]] <-
      ggplot(conc_data_id_splits[[j]], aes_string(x="TAFD", y=i)) +
      geom_point(shape=124, size = 1.5) +
      geom_line() +
      facet_wrap(~OSID+OSTUDYID, labeller = "label_both", nrow=3, ncol=4) +
      coord_cartesian(xlim = c(tafd_range$min, tafd_range$max)) +
      labs(y=i, x = labs_TAFD)
    }
  # join lists for all covariates
  time_var_cat_cov_plots <- c(time_var_cat_cov_plots, p1)
}

# takes some time to print if many individuals and many covariates...
invisible(lapply(time_var_cat_cov_plots, grob_draw))

if(params$print_results){
  pdf(file=file.path(res_eda_dir,
                     paste0("individual_time_varying_cat_cov_TAFD_", delivery_date,".pdf")),
      height=8.5, width=11)
  invisible(lapply(time_var_cat_cov_plots, grob_draw))
  dev.off()
}
```

