---
title: "Exploratory data analysis: Concentrations versus time"
author: "Author"
date: "Date"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
params:
  print_results: FALSE
---

```{r setup, include=F}
###################################################
# s05_eda_conc_time.Rmd
# Description: Exploratory data analysis of concentrations vs time for population PK analysis
# Dependencies: s01_datasetPrep.R / s01.RData
###################################################

# Settings to knit in top directory:
# Everything after this chunk works with paths relative top level
library(rprojroot)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 
knitr::opts_chunk$set(echo=F)

# Note: R markdown opens a new R session, your global environment is not available.

```

This script uses the data.frame "data", loaded from s01.RData. That is, all rows with C=="C" has been excluded for exploratory data analysis.
```{r environmentPrep, echo=T, message=F}
# -----------------------------------------------
# Prepare environment
# -----------------------------------------------
source(file = file.path("./Scripts","Setup","setup01_rEnvironment.R"))
load(file = file.path("./Scripts","s01.RData"))

```

Are the plots and tables also being written to file?
```{r}
params$print_results
```

# Numeric summaries
Number of observations in the dataset by BLQ.

```{r concentrationSummaries, echo=T}
conc_data %>% 
  group_by(BLQ) %>% 
  summarize(n = n()) %>% 
  mutate(percent = signif( 100 * n / sum(n), digits = 3))

```

Number of samples and occasions per subject (with/without BLQ)
```{r concentrationSummariesBySubject}
# total
conc_per_pat <- 
  conc_data %>%
  group_by(NMSEQSID) %>% 
  summarize(nDV=length(DV), 
            nOcc = ifelse(all(is.na(OCC)), 0, 
                          max(as.numeric(as.character(OCC)), na.rm=T)),
            maxTIME = max(TIME, na.rm=T)) %>% 
  ungroup() %>% 
  select(nDV, nOcc, maxTIME)

summary(conc_per_pat)

# >BLQ
conc_per_pat_noBLQ <-  
  conc_data %>%
  filter(BLQ == "Non-BLQ") %>% 
  group_by(NMSEQSID) %>% 
  summarize(nDV=length(DV), 
            nOcc = ifelse(all(is.na(OCC)), 0, 
                          max(as.numeric(as.character(OCC)), na.rm=T)),
            maxTIME = max(TIME, na.rm=T)) %>% 
  ungroup() %>% 
  select(nDV, nOcc, maxTIME)

summary(conc_per_pat_noBLQ)
```


Number of individuals and samples stratified by XXX

```{r samples, warning=F}
# The pre-writted code is stratifying on study. Can easily be changes to e.g. dose level

# Please note: if you are stratifying on DOSE and subjects 
# are gived multiple doses they are counted several times

# Including BLQ samples
conc_per_strata <-
  conc_data %>% 
  group_by(STUDYID) %>% 
  summarize(nSubjects = length(unique(NMSEQSID)),
            nConc = length(DV)) %>% 
  ungroup() %>% 
  mutate(pConcStrat = signif(100 * nConc / sum(nConc), digits=3))

# Excluding BLQ samples
conc_per_strata_noBLQ <-  
  conc_data %>% 
  filter(BLQ == "Non-BLQ") %>% 
  group_by(STUDYID) %>% 
  summarize(nSubjects = length(unique(NMSEQSID)),
            nConc = length(DV)) %>% 
  ungroup() %>% 
  mutate(pConcStrat = signif(100 * nConc / sum(nConc), digits=3))

# % BQL per strata
conc_per_strata <- 
  conc_per_strata %>% 
  bind_cols(nConc_NoBlq = conc_per_strata_noBLQ$nConc) %>% 
  mutate(pBlqByStrat = 
           signif(100 - (100 * nConc_NoBlq / nConc), digits=3)) %>% 
  select(STUDYID, nSubjects, nConc, pConcStrat, pBlqByStrat) %>% 
  arrange(rev(STUDYID)) %>% 
  rename(Study = STUDYID, 
         `Subjects (n)` = nSubjects, 
         `Concentrations (n)` = nConc, 
         `Concentrations (%)` = pConcStrat, 
         `BLQ (%)` = pBlqByStrat)

kable(conc_per_strata,
      caption = "Number of individuals and samples per study")

```

```{r, include=F}
# --------- Save to file
if(params$print_results){

  write.csv(conc_per_strata, 
            file = file.path(directories[["res_eda_dir"]], "conc_by_strat.csv"), 
            row.names = F)
  
}
```

# Samples vs time after dose

Add code for histograms of BLQ/non-blq histograms for TAD

# Stratified concentration versus time graphics

## Concentrations vs time after first dose
Lines connect data from one occasion within a subject. Colour indicate subject. Points indicate measured data. Dashed blue line show the lower limit of quantification. BLQ data is prited as LLOQ/2. 

```{r ConcTime, fig.height=6.5, fig.width=8.5, warning=F, message=F}
# Used to set the same x-axes in plots
range_axes <- conc_data %>% 
  summarize(minDV = min(DV, na.rm = T),
            maxDV = max(DV, na.rm = T), 
            maxTIME = max(TIME, na.rm = T),
            maxTAPD = max(TAPD, na.rm = T))

# List of plots
plots <- vector("list", length(conc_data_strat_split))

for(i in 1:length(conc_data_strat_split)){
  p <-
    arrangeGrob(
      # linear
      gg_conc_time(conc_data_strat_split[[i]], x=TIME, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, range_axes$maxTIME)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TIME, y=labs_conc),
      # log
      gg_conc_time(conc_data_strat_split[[i]], x=TIME, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, range_axes$maxTIME)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TIME, y=labs_conc) + 
        scale_y_log10(), 
      nrow=2)
  plots[[i]] <- p
}

# Print all the elements in the list on a separate page
walk(plots, grob_draw)


# Print to file
if(params$print_results){
  pdf(file = file.path(directories[["res_eda_dir"]],
                       paste0("conc_time_by_strat_", delivery_date,".pdf")),
      height=6.5, width=8.5)
  walk(plots, grob_draw)
  dev.off()
}
``` 

## Concentrations vs time after dose
*Right*: un-transformed scale. *Left*: log-transformed y-axis. Lines connect data from one occasion within a subject. Colour indicate subject. Points indicate measured data. Dashed blue line show the lower limit of quantification. BLQ data is prited as LLOQ/2. 

```{r ConcTad, fig.height=3, fig.width=8.5, warning=F, message=F}
# List of plots
plots <- vector("list", length(conc_data_strat_split))
for(i in 1:length(conc_data_strat_split)){
  p <-
    arrangeGrob(
      # linear
      gg_conc_time(conc_data_strat_split[[i]], x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, range_axes$maxTAPD)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc),
      # log
      gg_conc_time(conc_data_strat_split[[i]], x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, range_axes$maxTAPD)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc) + 
        scale_y_log10(), 
      nrow=1)
  plots[[i]] <- p
}


# Print all the elements in the list on a separate page
walk(plots, grob_draw)


# Print to file
if(params$print_results){
  pdf(file = file.path(directories[["res_eda_dir"]],
                       paste0("conc_vs_tad_by_strat_", delivery_date,".pdf")),
      height = 3, width = 8.5)
  walk(plots, grob_draw)
  dev.off()
}
```


## Concentrations versus time after dose (zoomed)

### First 12 hours
*Right*: un-transformed scale. *Left*: log-transformed y-axis. Lines connect data from one occasion within a subject. Colour indicate subject. Points indicate measured data. Dashed blue line show the lower limit of quantification. BLQ data is prited as LLOQ/2. 

```{r ConcTadZoom, fig.height=3, fig.width=8.5, warning=F, message=F} 
# If you want to zoom in on a specific time scale e.g. absorption phase or first 12/24 hr 
x_max <- 12

# List of plots
plots <- vector("list", length(conc_data_strat_split))

for(i in 1:length(conc_data_strat_split)){
  p <-
    arrangeGrob(
      # linear
      gg_conc_time(conc_data_strat_split[[i]], x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, x_max)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc),
      # log
      gg_conc_time(conc_data_strat_split[[i]], x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, x_max)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc) + 
        scale_y_log10(), 
      nrow=1)
  plots[[i]] <- p
}

# Print all the elements in the list on a separate page
walk(plots, grob_draw)

if(params$print_results){
  pdf(file=file.path(directories[["res_eda_dir"]],
                     paste0("conc_vs_tad_by_strat_", x_max, "_", delivery_date,".pdf")),
      height=3, width=8.5)
  walk(plots, grob_draw)
  dev.off()
}
```

### Absorption phase: first 4 hours
*Right*: un-transformed scale. *Left*: log-transformed y-axis. Lines connect data from one occasion within a subject. Colour indicate subject. Points indicate measured data. Dashed blue line show the lower limit of quantification. BLQ data is prited at LLOQ/2. 

```{r ConcTadAbs, fig.height=3, fig.width=8.5, warning=F, message=F} 
# Zoom in on absorption phase
x_max <- 4

# List of plots
plots <- vector("list", length(conc_data_strat_split))
for(i in 1:length(conc_data_strat_split)){
  p <-
    arrangeGrob(
      # linear
      gg_conc_time(conc_data_strat_split[[i]], x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, x_max)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc),
      # log
      gg_conc_time(conc_data_strat_split[[i]], x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT) + 
        coord_cartesian(xlim=c(0, x_max)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc) + 
        scale_y_log10(), 
      nrow=1)
  plots[[i]] <- p
}
# Print all the elements in the list on a separate page
walk(plots, grob_draw)

if(params$print_results){
  pdf(file=file.path(directories[["res_eda_dir"]], 
                     paste0("absorption_phase_", x_max,"_", delivery_date,".pdf")),
      height=3, width=8.5)
  walk(plots, grob_draw)
  dev.off()
}
```

## Comparison of single and multiple dose occasions
*Right*: un-transformed scale. *Left*: log-transformed y-axis. Lines connect data from one occasion within a subject. Colour indicate subject. Points indicate measured data. Dashed blue line show the lower limit of quantification. BLQ data is prited at LLOQ/2. 

```{r SingleVsMultiple, fig.height=6, fig.width=8.5, warning=F, message=F} 
# First 24 hours after dose, compare single and steady state
# (Requires that DOSEFLAG has been added to the dataset in s01_dataset_preparation.R)
x_max <- 24

# List of plots
plots <- vector("list", length(conc_data_strat_split))

for(i in 1:length(conc_data_strat_split)){
  # remove sparse occasions (DOSEFLAG = NA)
  dat <- conc_data_strat_split[[i]] %>% 
    filter(!is.na(DOSEFLAG))
  
  p <-
    arrangeGrob(
      # linear
      gg_conc_time(dat, x=TAPD,
                 y=DV, color=NMSEQSID, occ=OCC) + 
        facet_wrap(~STRATSPLIT+DOSEFLAG, nrow = 2) + 
        coord_cartesian(xlim=c(0, x_max)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc),
      # log
      gg_conc_time(dat, x=TAPD, 
                 y=DV, color=NMSEQSID, occ=OCC) +
        facet_wrap(~STRATSPLIT+DOSEFLAG, nrow = 2) + 
        coord_cartesian(xlim=c(0, x_max)) +
        guides(colour="none", shape="none") + 
        labs(x=labs_TAPD, y=labs_conc) + 
        scale_y_log10(), 
      nrow=1)
  plots[[i]] <- p
}

# Print all the elements in the list on a separate page
walk(plots, grob_draw)

# Print to file
if(params$print_results){ 
  pdf(file = file.path(directories[["res_eda_dir"]],
                       paste0("single_vs_multiple_dose_", x_max,"_", delivery_date,".pdf")),
      height=6, width=8.5)
  walk(plots, grob_draw)
  dev.off()
}
```




# Individual plots of concentrations vs time
The section below generates a list of individual plots with different combinations of the following settings:

* Concentration versus time after first dose *or* time after dose

* Log scale and untransformed scale

* Free/fixed scales for axes across individuals

Not all plots are going to be useful, you can just comment out/delete the ones you do not want to keep.


```{r individual, fig.height=8.5, fig.width=11, message=F, warning=F} 

# 1. Conc vs TAFD. Same axes across subjects. 
individualPlots <- list(
  gg_title_plot("Concentration vs. time after first dose \n Same axes across subjects"))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAFD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both",
               nrow=3, ncol=4) +
    # set scales for axes
    coord_cartesian(ylim=c(0, range_axes$maxDV), 
                    xlim=c(0, range_axes$maxTAFD)) +
    guides(shape="none") + 
    labs(x=labs_TAFD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 2. Log Conc vs. TAFD. Same axes across subjects
individualPlots <- 
  c(individualPlots, list(
    gg_title_plot("Concentrations vs. time after first dose \n Same axes across subjects \n\n Semi-log")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAFD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both", 
               nrow=3, ncol=4) +
    # set scales for axes
    coord_cartesian(ylim=c(range_axes$minDV, range_axes$maxDV), 
                    xlim=c(0, range_axes$maxTAFD)) +
    scale_y_log10(breaks=c(1,10,100,1000)) +
    guides(shape="none") + 
    labs(x=labs_TAFD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 3. Conc vs TAFD. Free scales on both axes
individualPlots <-
  c(individualPlots, list(
    gg_title_plot("Concentration. vs time after first dose \n Free scales on axes")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAFD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both", 
               nrow=3, ncol=4, 
               scales="free") +
    guides(shape="none") + 
    labs(x=labs_TAFD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 4. Log Conc vs TAFD. Free scales on axes
individualPlots <-
  c(individualPlots,list(
    gg_title_plot("Concentration vs. time after first dose \n Free scales on axes \n\n Semi-log")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAFD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both", 
               nrow=3, ncol=4, 
               scales="free") +
    scale_y_log10(breaks=c(1,10,100,1000)) +
    guides(shape="none") + 
    labs(x=labs_TAFD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 5. Conc vs. TAPD. Same y-axis across subjects. 
individualPlots <- 
  c(individualPlots, list(
    gg_title_plot("Concentration vs. time after dose \n Same axes across subjects")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAPD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both",
               nrow=3, ncol=4) +
    # set scales for axes
    coord_cartesian(ylim=c(0, range_axes$maxDV), 
                    xlim=c(0, range_axes$maxTAPD)) +
    guides(shape="none") + 
    labs(x=labs_TAPD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 6. Log Conc vs. TAPD. Same y-axis across subjects
individualPlots <-
  c(individualPlots,list(
    gg_title_plot("Concentration vs. time after dose \n Same axes across subjects \n\n Semi-log")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAPD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both",
               nrow=3, ncol=4) +
    # set scales for axes
    coord_cartesian(ylim=c(range_axes$minDV, range_axes$maxDV),
                    xlim=c(0, range_axes$maxTAPD)) +
    scale_y_log10(breaks=c(1,10,100,1000)) +
    guides(shape="none") + 
    labs(x=labs_TAPD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 7. Conc vs TAPD - Free scales
individualPlots <-
  c(individualPlots,list(
    gg_title_plot("Concentration vs. time after dose \n Free scales on axes")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAPD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both",
               nrow=3, ncol=4, 
               scales="free") +
    guides(shape="none") + 
    labs(x=labs_TAPD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

## 8. Log conc vs TAPD - Free scales
individualPlots <-
  c(individualPlots,list(
    gg_title_plot("Concentration vs. time after dose \n Free scales on axes \n Semi-log")))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAPD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both",
               nrow=3, ncol=4, 
               scales="free") +
    scale_y_log10(breaks=c(1,10,100,1000)) +
    guides(shape="none") + 
    labs(x=labs_TAPD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

# 9. Log Conc vs. TAPD. Same y-axis across subjects, TAD <12 h
x_max <- 12

individualPlots <-
  c(individualPlots,list(
    gg_title_plot(paste("Concentration vs. time after dose \n TAD <", x_max ,"\n\n Semi-log"))))

p1 <- vector("list", length(conc_data_id_splits))
for(i in 1:(length(conc_data_id_splits))){
  p <-
    gg_conc_time(conc_data_id_splits[[i]], x=TAPD, y=DV, 
               color=REGIMEN, occ=OCC) + 
    facet_wrap(~OSID+COHORT, labeller="label_both",
               nrow=3, ncol=4) +
    # set scales for axes
    coord_cartesian(ylim=c(range_axes$minDV, range_axes$maxDV), 
                    xlim=c(0, x_max)) +
    scale_y_log10(breaks=c(1,10,100,1000)) +
    guides(shape="none") + 
    labs(x=labs_TAPD, y=labs_conc)
  p1[[i]] <- p
}
individualPlots <- c(individualPlots, p1)

## This list takes quite some time to print...
# Print all the elements in the list on a separate page
walk(individualPlots, grob_draw)


if(params$print_results){
  pdf(file = file.path(directories[["res_eda_dir"]],
                       paste0("individual_conc_vs_time_", delivery_date,".pdf")),
      height=8.5, width=11)
  walk(individualPlots, grob_draw)
  dev.off()
}
```
