---
title: "Exploratory data analysis: PK Linearity"
author: "Author"
date: ""
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
params:
  printResults: FALSE
---
```{r setup, include=F}
###################################################
# s06_EDAPkLinearity.Rmd
# Description: Exploratory data analysis of PK linearity for population PK analysis
# Dependencies: s01_datasetPrep.R
###################################################
library(rprojroot)
knitr::opts_chunk$set(echo=F)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 
# for correct printing of dust tables
options(pixie_interactive=F)

# Note: R markdown opens a new R session, your global environment is not available.
```

This script uses the data.frame "data", loaded from s01.RData. That is, all rows with C=="C" has been excluded for exploratory data analysis.
```{r environmentPrep, echo=T, message=F}
# ------------------------------------------------------------------
#  Prepare environment
# ------------------------------------------------------------------
# load packages
source(file=file.path("Scripts","Setup","setup01_projectPackages.R")) 
# load output from s01_DatasetPrep.R
load(file=file.path("Scripts","s01.RData"))

# White background in plots
theme_set(theme_bw()) # to be replaced with a azTheme
update_geom_defaults("point", list(shape = 1))
```

Are the plots and tables also being written to file?
```{r}
params$printResults
```

# Included data
```{r, include=F} 
# The sections below will need to be tweaked for each specific project. 
# This should only serve as an example

# ------------------------------------------------------------------
#  Settings for NCA:
# ------------------------------------------------------------------
ncaIntervals <-
  data.frame(start     = 0,
             end       = c(24),
             cmax      = T,
             tmax      = T,
             aucinf.obs= T,
             auclast   = T,
             tlast     = T,
             tfirst    = T,
             half.life = T,
             lambda.z  = T,
             lambda.z.n.points = T)

# ------------------------------------------------------------------
#  Extract single dose data
# ------------------------------------------------------------------
ncaSingle <-
  data %>% 
  filter(!is.na(DOSEFLAG) & DOSEFLAG =="Single dose")

# Use ID.OCC as ID to include multiple single doses from the same patient
# ID.OCC are missing for doses => fill backwards
ncaSingle <-
  ncaSingle %>% 
  mutate(ID.OCC = na.locf(ID.OCC, fromLast = T), 
         ID.OCC = as.numeric(as.character(ID.OCC)))

# Dose dataset
ncaSingleDose <- 
  ncaSingle %>% 
  filter(EVID==1) # dosing events
# unique(ncaSingle$AMT)

## Concentration dataset
ncaSingleConc <- 
  ncaSingle %>% 
  #  filter(EVID !=1) %>% # no dosing events
  filter(!is.na(BLQ) & BLQ=="Non-BLQ" & !is.na(DV)) # remove all BLQs

## You may have to remove other missing concentrations (MDV==1)
# ncaSingleConc <- 
#   ncaSingleConc %>% 
#   filter(!(!is.na(MDV) & MDV==1 & is.na(BLQ)))

# Add conc=0 at time 0 (or whatever you think is appropriate)
ncaSingleConc0 <- 
  data.frame(ID.OCC = unique(ncaSingleConc$ID.OCC),
             TAFD=0, 
             TAPD=0, 
             DV=0)

# join and sort
ncaSingleConc <- 
  full_join(ncaSingleConc, ncaSingleConc0)
ncaSingleConc <- 
  ncaSingleConc[order(ncaSingleConc$ID.OCC, ncaSingleConc$TAFD),]
```

## Plots of occasions included for single dose NCA parameters
Lines connect data from one occasion, points show measured concentrations, the blue horizontal line indicated time of dosing. 

```{r, fig.height=8.5, fig.width=11, warning=F, message=F}
# ------------------------------------------------------------------
#  Plot single dose data
# ------------------------------------------------------------------
# split by id.occ
ncaSingleDoseSplit <- ind_data_split(ncaSingleDose, 
                                   n_per_page = 12, id="ID.OCC")
ncaSingleConcSplit <- ind_data_split(ncaSingleConc,
                                   n_per_page = 12, id="ID.OCC")

# IndividualPK profiles for single dose
p1 <- vector("list", length(ncaSingleConcSplit))
for(i in 1:length(ncaSingleConcSplit)){
  p <-
    # concentrations
    ggplot(data=ncaSingleConcSplit[[i]], aes(x=TAPD, y=DV)) +
    geom_line() + geom_point() +
    # doses
    geom_vline(data = ncaSingleDoseSplit[[i]], aes(xintercept=TAPD),
               col="blue", linetype="dashed") +
    facet_wrap(~ID.OCC, nrow=3, ncol=4,
               scales="free_y", labeller="label_both") +
    labs(x=labTAPD, y=labConc)
  p1[[i]] <- p
}

invisible(lapply(p1, grob_draw))
```

```{r, include=F}
if(params$printResults){
  pdf(file=file.path(resEDADir, paste0("ncaSingleDoseData_", deliveryDate,".pdf")),
      height=8.5, width=11)
  invisible(lapply(p1, grob_draw))
  dev.off()
}
```

```{r, include=F}
# ------------------------------------------------------------------
# Merge conc and dose data into NCA object
# ------------------------------------------------------------------
ncaSingleDat <-
  PKNCAdata(ncaSingleConc, ncaSingleDose,
            formula.conc = DV~TAPD | ID.OCC,
            formula.dose = AMT~TAPD | ID.OCC,
            intervals = ncaIntervals)

# ------------------------------------------------------------------
# Run NCA, extract and re-format results
# ------------------------------------------------------------------
ncaResultsSingle <- pk.nca(ncaSingleDat)

# nca results
ncaResultsSingleExtract <- ncaResultsSingle$result
ncaResultsSingleExtract <- 
  spread(ncaResultsSingleExtract[, c("ID.OCC","PPTESTCD","PPORRES")],
         key = PPTESTCD, value = PPORRES)

# Merge back with covariate data (extract from dose object)
covExtractSingle <- 
  ncaResultsSingle$data$dose$data[, c("ID.OCC", contCov, bContCov, catCov, bCatCov,
                                      "OSID","OSTUDYID","AMT","FREQ", "DOSEFLAG")]

# unique(unique(covExtractSingle$ID.OCC) == unique(ncaResultsSingleExtract$ID.OCC))
resultsSingle <- full_join(ncaResultsSingleExtract, 
                           covExtractSingle)

names(resultsSingle)[names(resultsSingle)=="AMT"] <- "dose"
# unique(resultsSingle$Dose)
```

```{r, include=F}
# ------------------------------------------------------------------
#  Extract multiple dose/steady state data
# ------------------------------------------------------------------
ncaMultiple <-
  data %>% 
  filter(!is.na(DOSEFLAG) & DOSEFLAG =="Multiple dose")

# Use ID.OCC as ID to include multiple occasions from the same patient
# ID.OCC are missing for doses => fill backwards
ncaMultiple <-
  ncaMultiple %>% 
  mutate(ID.OCC = na.locf(ID.OCC, fromLast = T), 
         ID.OCC = as.numeric(as.character(ID.OCC)))

# Dose dataset
ncaMultipleDose <- 
  ncaMultiple %>% 
  filter(EVID==1) # dosing events
# unique(ncaMultiple$AMT)

## Concentration dataset
ncaMultipleConc <- 
  ncaMultiple %>% 
  #  filter(EVID !=1) %>% # no dosing events
  filter(!is.na(BLQ) & BLQ=="Non-BLQ" & !is.na(DV)) # remove all BLQs

## You may have to remove other missing concentrations (MDV==1)
# ncaMultipleConc <- 
#   ncaMultipleConc %>% 
#   filter(!(!is.na(MDV) & MDV==1 & is.na(BLQ)))

# Assume no accumulation and add a 0 conc at TAPD=0
# (or whatever you think is appropriate in your case)
ncaMultipleConc0 <- 
  data.frame(ID.OCC = unique(ncaMultipleConc$ID.OCC),
             TAFD=0, 
             TAPD=0, 
             DV=0)

# join and sort
ncaMultipleConc <- 
  full_join(ncaMultipleConc, ncaMultipleConc0)
ncaMultipleConc <- 
  ncaMultipleConc[order(ncaMultipleConc$ID.OCC, ncaMultipleConc$TAFD),]
```

## Plots of occasions included for multiple dose/steady state NCA parameters
Lines connect data from one occasion, points show measured concentrations, the blue horizontal line indicated time of dosing. 

```{r, fig.height=8.5, fig.width=11, warning=F, message=F}
# ------------------------------------------------------------------
#  Plot multiple dose data
# ------------------------------------------------------------------
# split by id.occ
ncaMultipleDoseSplit <- ind_data_split(ncaMultipleDose, 
                                     n_per_page = 12, id="ID.OCC")
ncaMultipleConcSplit <- ind_data_split(ncaMultipleConc,
                                     n_per_page = 12, id="ID.OCC")

# IndividualPK profiles for multiple dose/ steady state
p1 <- vector("list", length(ncaMultipleConcSplit))
for(i in 1:length(ncaMultipleConcSplit)){
  p <-
    # concentrations
    ggplot(data=ncaMultipleConcSplit[[i]], aes(x=TAPD, y=DV)) +
    geom_line() + geom_point() +
    # doses
    geom_vline(data = ncaMultipleDoseSplit[[i]], aes(xintercept=TAPD),
               col="blue", linetype="dashed") +
    facet_wrap(~ID.OCC, nrow=3, ncol=4,
               scales="free_y", labeller="label_both") +
    labs(x=labTAPD, y=labConc)
  p1[[i]] <- p
}

invisible(lapply(p1, grob_draw))
```

```{r, include=F}
if(params$printResults) {
  pdf(file=file.path(resEDADir, paste0("ncaMultipleDoseData_",deliveryDate,".pdf")),
      height=8.5, width=11)
  invisible(lapply(p1, grob_draw))
  dev.off()
}
```

```{r, include=F}
# ------------------------------------------------------------------
# Merge conc and dose data into NCA object
# ------------------------------------------------------------------
ncaMultipleDat <-
  PKNCAdata(ncaMultipleConc, ncaMultipleDose,
            formula.conc = DV~TAPD | ID.OCC,
            formula.dose = AMT~TAPD | ID.OCC,
            intervals = ncaIntervals)

# ------------------------------------------------------------------
# Run NCA, extract and re-format results
# ------------------------------------------------------------------
ncaResultsMultiple <- pk.nca(ncaMultipleDat)

# nca results
ncaResultsMultipleExtract <- ncaResultsMultiple$result
ncaResultsMultipleExtract <- 
  spread(ncaResultsMultipleExtract[, c("ID.OCC","PPTESTCD","PPORRES")],
         key = PPTESTCD, value = PPORRES)

# Merge back with covariate data (extract from dose object)
covExtractMultiple <- 
  ncaResultsMultiple$data$dose$data[, c("ID.OCC", contCov, bContCov, catCov, bCatCov,
                                        "OSID","OSTUDYID","AMT","FREQ", "DOSEFLAG")]

# unique(unique(covExtractMultiple$ID.OCC) == unique(ncaResultsMultipleExtract$ID.OCC))
resultsMultiple <- full_join(ncaResultsMultipleExtract, 
                             covExtractMultiple)

names(resultsMultiple)[names(resultsMultiple)=="AMT"] <- "dose"
# unique(resultsMultiple$Dose)

# ------------------------------------------------------------------
#  Merge restults from sinlge dose and multiple dose occasions
# ------------------------------------------------------------------
results <- rbind(resultsSingle, resultsMultiple)

### Calculate dose normalized values
results$aucInfDoseNorm <- (results$aucinf.obs / results$dose)
results$aucLastDoseNorm <- (results$auclast / results$dose)
results$cmaxDoseNorm   <- (results$cmax / results$dose)
```

# Assessment of PK linearity, $AUC_{Inf}$ vs dose
## ANOVA
Add figure text here

```{r, fig.height=5, fig.width=6.5, warning=F, message=F}
anova <- dose_proportionality_anova(results, x=dose, y=aucinf.obs, 
                                    plot.par = T)
anova[["plot"]] + 
  labs(y=expression(AUC[inf]~(mg/L%.%h)), 
       x=expression(Dose~(mg)))
```

## Linear model
Add figure text here

```{r, fig.height=5, fig.width=6.5, warning=F, message=F}
linear <- dose_proportionality_linear(results, x=dose, y=aucinf.obs, 
                                      plot.par = T)
linear[["plot"]] + 
  labs(y=expression(AUC[inf]~(mg/L%.%h)), 
       x=expression(Dose~(mg)))
```

## Power-model
Add figure text here

```{r, fig.height=5, fig.width=6.5, warning=F, message=F}
power <- dose_proportionality_power(results, x=dose, y=aucinf.obs, 
                                      plot.par = T)
power[["plot"]] + 
  labs(y=expression(AUC[inf]~(mg/L%.%h)), 
       x=expression(Dose~(mg)))
```

# Other graphics assessing PK linearity
Add figure texts here

```{r, fig.height=4.5, fig.width=5.5, warning=F, message=F}
# Dose should be numeric (it is set as a group within boxplot) using gg_continous_box
# That way the x-axis is still numeric and displayes the data as continous
# (important if large range of doses with long, not equally long steps between)
# str(results$dose)

linearityPlotsPool <- vector("list", 0)

## Cmax ###
linearityPlotsPool$cmax <-
  gg_box_continuous(results, x=dose, y=cmax) + 
  labs(x=expression(Dose~(mg)), y=expression(C[max]~(mg/L)))

linearityPlotsPool$normCmax <-
  gg_box_continuous(results, x=dose, y=cmaxDoseNorm) + 
  geom_hline(aes(yintercept=mean(cmaxDoseNorm, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression(Dose~normalized~C[max]~(mg/L)),
       x = expression(Dose~(mg)))

## AUC Infinity ###
linearityPlotsPool$aucInf <-
  gg_box_continuous(results, x=dose, y=aucinf.obs) + 
  labs(y=expression(AUC[inf]~(mg/L%.%h)), 
       x=expression(Dose~(mg)))

linearityPlotsPool$normAucInf <-
  gg_box_continuous(results, x=dose, y=aucInfDoseNorm) + 
  geom_hline(aes(yintercept=mean(aucInfDoseNorm, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression(Dose~normalized~AUC[inf]~(mg/L%.%h)),
       x = expression(Dose~(mg)))

## AUC 24h ###
linearityPlotsPool$aucLast <-
  gg_box_continuous(results, x=dose, y=auclast) + 
  labs(y=expression(AUC[last]~(mg/L%.%h)), 
       x = expression(Dose~(mg)))

linearityPlotsPool$NormAUCinf <-
  gg_box_continuous(results, x=dose, y=aucLastDoseNorm) + 
  geom_hline(aes(yintercept=mean(aucLastDoseNorm, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression(Dose~normalized~AUC[last]~(mg/L%.%h)), 
       x = expression(Dose~(mg)))

# t1/2 ##
linearityPlotsPool$t_half <-
  gg_box_continuous(results, x=dose, y=half.life) + 
  geom_hline(aes(yintercept=mean(half.life, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression("t"["1/2"]~(h)),
       x = expression(Dose~(mg)))

invisible(lapply(linearityPlotsPool, grob_draw))

if(params$printResults){
  pdf(file=file.path(resEDADir,
                     paste0("PKLinearityAssessment_pooled_", deliveryDate,".pdf")),
      height=4.5, width=5.5)
  invisible(lapply(linearityPlotsPool, grob_draw))
  dev.off()
}
```

## Stratified by single/multiple dose
Add figure text here

```{r, fig.height=5, fig.width=7.5, warning=F, message=F}
linearityPlotsSplit <- vector("list", 0)

## Cmax ###
linearityPlotsSplit$CmaxSplit <-
  gg_box_continuous(results, y=cmax, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  labs(y=expression(C[max]~(mg/L)),
       x=expression(Dose~(mg)))

linearityPlotsSplit$normCmaxSplit <- 
  gg_box_continuous(results, y=cmaxDoseNorm, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  # add grand mean
  geom_hline(aes(yintercept=mean(cmaxDoseNorm, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression(Dose~normalized~C[max]~(mg/L)),
       x=expression(Dose~(mg)))

## AUC Infinity ###
linearityPlotsSplit$aucInfSplit <-
  gg_box_continuous(results, y=aucinf.obs, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  labs(y=expression(AUC[inf]~(mg/L%.%h)),
       x=expression(Dose~(mg)))

linearityPlotsSplit$normAucInfSplit <-
  gg_box_continuous(results, y=aucInfDoseNorm, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  # add grand mean
  geom_hline(aes(yintercept=mean(aucInfDoseNorm, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression(Dose~normalized~AUC[inf]~(mg/L%.%h)),
       x = expression(Dose~(mg)))

## AUC Last ###
linearityPlotsSplit$AUCinfSplit <-
  gg_box_continuous(results, y=auclast, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  labs(y=expression(AUC[last]~(mg/L%.%h)),
       x=expression(Dose~(mg)))

linearityPlotsSplit$NormAUCinfSplit <-
  gg_box_continuous(results, y=auclast, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  # add grand mean
  geom_hline(aes(yintercept=mean(aucLastDoseNorm, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression(Dose~normalized~AUC[last]~(mg/L%.%h)),
       x = expression(Dose~(mg)))

## t1/2 ##
linearityPlotsSplit$t_halfSplit <-
  gg_box_continuous(results, y=half.life, x=dose,
                  fill=DOSEFLAG, dodge=30) +
  # add grand mean
  geom_hline(aes(yintercept=mean(half.life, na.rm = T)),
             linetype="dashed", size=0.6) +
  labs(y=expression("t"["1/2"]~(h)),
       x=expression(Dose~(mg)))

invisible(lapply(linearityPlotsSplit, grob_draw))

if(params$printResults){
  pdf(file=file.path(resEDADir,
                     paste0("PKLinearityAssessment_stratified_", deliveryDate,".pdf")),
      height=8.5, width=11)
  invisible(lapply(linearityPlotsSplit, grob_draw))
  dev.off()
}
```

# Assessment of covariate effect on NCA parameters
## Continuous covariates
Add figure text here

```{r, fig.height=8, fig.width=10, warning=F, message=F}
doseNormVsCont <-
  ggduo(results, bContCov, c("aucInfDoseNorm",
                             "aucLastDoseNorm",
                             "cmaxDoseNorm"),
        types = list(continuous = "smooth_loess"),
        columnLabelsY = c("AUC[Inf]/Dose~(mg/L%.%h/mg)",
                          "AUC[last]/Dose~(mg/L%.%h/mg)",
                          "C[max]/Dose~(mg/L/mg)"),
        labeller="label_parsed")

doseNormVsCont

if(params$printResults){
  pdf(file=file.path(resEDADir,
                     paste0("NCA_ContinousCovariates", deliveryDate,".pdf")),
      height=8, width=10)
  print(doseNormVsCont)
  dev.off()
}
```

## Categorical covariates
Add figure text here

```{r, fig.height=8, fig.width=10, warning=F, message=F}
doseNormVsCat <-
  ggduo(results, bCatCov, c("aucLastDoseNorm",
                            "aucLastDoseNorm",
                            "cmaxDoseNorm"),
        types = list(comboVertical = ally_box),
        columnLabelsY = c("AUC[Inf]/Dose~(mg/L%.%h/mg)",
                          "AUC[Last]/Dose~(mg/L%.%h/mg)",
                          "C[max]/Dose~(mg/L/mg)"),
        labeller="label_parsed")

doseNormVsCat

if(params$printResults){
  pdf(file=file.path(resEDADir,
                 paste0("NCA_CategoricalCovariates", deliveryDate,".pdf")),
      height=8, width=10)
  print(doseNormVsCat)
  dev.off()
}
```

## Stratified by study
Add figure text here

```{r, fig.height=8, fig.width=5, warning=F, message=F}
doseNormVsStudy <-
  ggduo(results,"OSTUDYID", c("aucLastDoseNorm",
                            "aucLastDoseNorm",
                            "cmaxDoseNorm"),
        types = list(comboVertical = ally_box),
        columnLabelsY = c("AUC[Inf]/Dose~(mg/L%.%h/mg)",
                          "AUC[Last]/Dose~(mg/L%.%h/mg)",
                          "C[max]/Dose~(mg/L/mg)"),
        labeller="label_parsed")

if(params$printResults){
  pdf(file=file.path(resEDADir,
                 paste0("NCA_ByStudy", deliveryDate,".pdf")),
      height=8, width=5)
  print(doseNormVsStudy)
  dev.off()
}
```

```{r, include=F}
## Empty environment
# rm(list = ls())
```