---
title: "NONMEM dataset preparation"
author: "Author"
date: "Date"
output: 
  html_document:
params:
  printCSV: FALSE
---

```{r setup, include=F}
###################################################
# s07_NMDatasets.Rmd
# Description: Dataset preparation for NONMEM
# Dependencies: s01_datasetPrep.R 
###################################################
library(rprojroot)
knitr::opts_chunk$set(echo=T)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 

# Note: R markdown opens a new R session, your global environment is not available.
```

This script uses the data.frame "rawdata". That is, all rows with C=="C" are included (as required by regulatory agencies and will be excluded in the NONMEM scripts. 
```{r environmentPrep, message=F}
# ------------------------------------------------------------------
#  Prepare environment
# ------------------------------------------------------------------
# load packages
source(file=file.path("Scripts","Setup","setup01_projectPackages.R")) 
# load output from s01_DatasetPrep.R
load(file=file.path("Scripts","s01.RData"))
```

Are the datasets being outputted to file?
```{r, echo=F}
params$printCSV
```

#  Selection of columns and re-naming for NONMEM 
```{r}
# names(rawdata)

# Only numeric columns
cols <- c("C","NMSEQSID","TAFD","TAPD","AMT","DV","LNDV","MDV","EVID",
          "CMT","BLQ","OCC","STUDYID","COHORT","DOSE","AGE","SEXM","RACE",
          "ETHNIC","BSCR","BEGFR","BRENAL","BHT","BWT","BBMI","FREQ")

nmData <- rawdata[ , cols]  # use rawdata: contains all "C" and numeric versions of columns

# Comment out rows with negative time (pre-first dose samples)
nmData$C[nmData$TAFD < 0] <- "C"

# rename NMSEQSID to ID and TAFD to TIME
names(nmData)[names(nmData)=="NMSEQSID"] <- "ID"
names(nmData)[names(nmData)=="TAFD"] <- "TIME"

# # use ln transformed data
# names(nmData)[names(nmData)=="DV"] <- "NORMDV"
# names(nmData)[names(nmData)=="LNDV"] <- "DV"
```

# Imputation of missing covariates
```{r}
baselineData <- nmData[!duplicated(nmData$ID), ]
# summary(baselineData) 

## Categorical covariates: model as separate groups and potentially merge with other
nmData$BRENAL[is.na(nmData$BRENAL)]          <- -99

## Continous covariates: impute based on median (stratified by sex if appropriate)
# Height 
# nmData$BHT[is.na(nmData$BHT) & nmData$SEXM==0 ] <-
#   median(baselineData$BHT[baselineData$SEXM==0], na.rm=T)
nmData$BHT[is.na(nmData$BHT) & nmData$SEXM==1 ] <- 
  median(baselineData$BHT[baselineData$SEXM==1], na.rm=T)

# Weight 
# nmData$BWT[is.na(nmData$BWT) & nmData$SEXM==0 ] <-
#   median(baselineData$BWT[baselineData$SEXM==0], na.rm=T)
nmData$BWT[is.na(nmData$BWT) & nmData$SEXM==1 ] <-
  median(baselineData$BWT[baselineData$SEXM==1], na.rm=T)

# Calculate BMI based on imputed (and reported) values
nmData$BBMI[is.na(nmData$BBMI)] <-
  nmData$BWT[is.na(nmData$BBMI)] / (nmData$BHT[is.na(nmData$BBMI)]/100)^2

# # Calculate BSA based on imputed (and reported) values (Dubois&Dubois)
# nmData$BBSA[is.na(nmData$BBSA)] <- 
#   0.007184 * ((nmData$BHT[is.na(nmData$BBSA)])^0.725) * (nmData$BWT[is.na(nmData$BBSA)]^0.425)

# # CrCL
# nmData$BCRCL[is.na(nmData$BCRCL) & nmData$SEXM==0] <- 
#   median(baselineData$BCRCL[baselineData$SEXM==0], na.rm=T)
# nmData$BCRCL[is.na(nmData$BCRCL) & nmData$SEXM==1] <- 
#   median(baselineData$BCRCL[baselineData$SEXM==1], na.rm=T)

# eGFR
nmData$BEGFR[is.na(nmData$BEGFR) & nmData$SEXM==0] <-
  median(baselineData$BEGFR[baselineData$SEXM==0], na.rm=T)
nmData$BEGFR[is.na(nmData$BEGFR) & nmData$SEXM==1] <- 
  median(baselineData$BEGFR[baselineData$SEXM==1], na.rm=T)

# SCR
nmData$BSCR[is.na(nmData$BSCR)] <- 
  median(baselineData$BSCR, na.rm=T)
```

# Output and naming of dataset
The dataset names are saved in variables that are used to update the nonmem control-files later on.

## No. 1: original with imputed covariates
```{r}
# Dataset name and path to location
nmDataName <- paste0(drugName,"_NM_", deliveryDate,".csv")
outFile <- file.path(derivedDataDir, nmDataName)

# Write dataset 
if(params$printCSV){ 
  write.csv(nmData, file=outFile, row.names=F, quote=F, na = ".") 
}
```

## No. 2: added RATE for estimation of zero-order input
```{r}
## Add a RATE column with values -2 at dosing records 
## for estimation of zero order input (for models with zero order input)
nmDataComb <- nmData
nmDataComb$RATE <- rep(NA, nrow(nmDataComb))
nmDataComb$RATE[!is.na(nmDataComb$AMT)] <- -2

# Dataset name and path to location
nmDataNameComb <- paste0(drugName, "_NM_", deliveryDate, "_comb",".csv")
outFile <- file.path(derivedDataDir, nmDataNameComb)
  
# Write dataset
if(params$printCSV){
  write.csv(nmDataComb, file=outFile, row.names=F, quote=F, na = ".")
}
```

## No. 3: added duplicated dose records for estimation of sequential or parallell absorption
```{r}
nmDataPar <- nmDataComb   # need the RATE column so use nmDataComb

# Duplicate all dose records for estimation of zero-order input into central (cmt==2)
doses <- nmDataPar[nmDataPar$EVID==1,] 
doses$CMT <- 2

# Modify record in dataset for which we use Ka
nmDataPar$RATE[nmDataPar$EVID==1] <- NA

# Merge duplicated doses with data and sort by ID and time
nmDataPar <- rbind(nmDataPar, doses)
nmDataPar <- nmDataPar[order(nmDataPar$ID, nmDataPar$TIME), ]

# Dataset name and path to location
nmDataNamePar <- paste0(drugName, "_NM_", deliveryDate, "_par",".csv")
outFile <- file.path(derivedDataDir, nmDataNamePar)

# Write dataset
if(params$printCSV){
  write.csv(nmDataPar, file=outFile, row.names=F, quote=F, na = ".")
}
```

## No. 4: added dummy rows for multiple dose transit compartment model
```{r}
## Add dummy variables at time of dosing in order 
## to aviod negative TAD for transit compartment model
nmDataTransit <- nmData

# Use dose records and modify to merge right before next dose
transitDummy <- nmDataTransit[!is.na(nmDataTransit$AMT) & 
                                nmDataTransit$TIME !=0, ]  # not needed for first dose

# change evid to 2, the time to 10 min before the dose, AMT to 0
# only baseline values for covariates so does not need to change these
transitDummy$EVID <- 2
transitDummy$AMT <- NA
transitDummy$TIME <- transitDummy$TIME - (10/60)
transitDummy$TAPD <- NA
transitDummy$CMT <- 2

# Bind with the original data and sort by ID and time
nmDataTransit <- rbind(nmDataTransit, transitDummy)
nmDataTransit <- nmDataTransit[order(nmDataTransit$ID, nmDataTransit$TIME),]
rownames(nmDataTransit) <- NULL

# Dataset name and path to location
nmDataNameTransit <- paste0(drugName, "_NM_", deliveryDate, "_transit",".csv")
outFile <- file.path(derivedDataDir, nmDataNameTransit)

# Write dataset 
if(params$printCSV){
  write.csv(nmDataTransit, file=outFile, row.names=F, quote=F, na = ".")
}
```

# List of prepared datasets
```{r, echo=F}
# Save dataframe with number of datasets and modifications done. 
NMdatasets <- 
  data.frame(FileName = c(nmDataName,nmDataNameComb,nmDataNamePar,nmDataNameTransit), 
            Description = 
              c("Dataset prepared for NONMEM use: (i) only numerical columns selected, (ii) missing covariates imputed, (iii) NMSEQSID renamed to ID, (iv) TAFD renamed to TIME",
                "RATE column added to estimate duration of zero-order absorption", 
                "Duplicated dose record added for estimation of sequential or parallel zero- and first order absorption",
                "Dummy variable added at 10 min before each dose in order to avoid negative time after dose results for transit compartment model"))
dust(NMdatasets)
```

```{r, echo=F}
# Write dataset 
if(params$printCSV){
  write.csv(NMdatasets, 
            file=file.path(derivedDataDir, "preparedDatasets.csv"), 
            row.names=F, quote=T)
}
```

# Save environment for use in model development
```{r}
# Save environment to be used in developement scripts 
rm("params") # remove rmarkdown params or it will cause issues in the next script
# this will generate in error but all the output is generated as should.
save.image(file=file.path(scriptsDir,"s07.RData"))
```
