---
title: "Summary of excluded data"
author: "Author"
date: "Date"
output: html_document
params:
  print_results: FALSE
---

```{r setup, include=FALSE}
###################################################
# s03_summary_excluded_data.R
# Description: Summary of data excluded with C="C"
# Dependencies: s01_dataset_preparation.R / s01.RData
###################################################

# Settings to knit in top directory:
# Everything after this chunk works with paths relative top level
library(rprojroot)
knitr::opts_knit$set(root.dir=find_root(has_file("OpenProject.Rproj"))) 
knitr::opts_chunk$set(echo=F)

# Note: R markdown opens a new R session, your global environment is not available.
```

Load the dataset as prepared by "s01_dataset_preparation.R"
```{r, echo=T, message=F}
# -----------------------------------------------
# Prepare environment
# -----------------------------------------------
source(file = file.path("./Scripts","Setup","setup01_rEnvironment.R"))
load(file = file.path("./Scripts","s01.RData"))

```

The outputs from this scripts are: 

  1. Summary table of number of excluded data and reasons (split by >/< LLOQ)
  
  2. Table listing each excluded record
  
Both are intended to be used in the report (appendices)


# Summary of excluded subjects, concentrations and doses
Number of subjects excluded and total number in final data:
```{r}
# data frame 'rawdata' contains all rows. 
# In the data frame 'data' rows with c="C" has been removed
rawdata %>% summarize(nSubTot = length(unique(NMSEQSID)))

data    %>% summarize(nSubIncl = length(unique(NMSEQSID)))
```

Total number of excluded rows: 
```{r}
excluded <- rawdata %>% filter(C=="C")
nrow(excluded)

excl_doses <- excluded %>% filter(EVID==1)
excl_conc  <- excluded %>% filter(EVID==0)
```

Out of which `r nrow(excl_doses)` rows are doses and `r nrow(excl_conc)` concentrations.

Is anything else than EVID=1 or EVID=0 that is excluded?
```{r}
test <- sum(nrow(excl_doses), nrow(excl_conc)) !=  nrow(excluded)

check_message(test)
```

How much of the excluded concentration data is missing or BLQ?
```{r, echo=T}
excl_conc_missing <- 
  excl_conc %>% filter(MDV==1)

nrow(excl_conc_missing) # rows
100*( nrow(excl_conc_missing) / nrow(excl_conc) )  # percent
```


# Summarize excluded data

## Review reasons for exclusion
### Excluded doses
```{r}
unique(excl_doses[,"COMMENT"])
```

### Excluded concentrations
```{r}
unique(excl_conc[,"COMMENT"])
```

Primary reason for exclusion split by BLQ/non-BLQ:
```{r}
# In this dataset there are not multiple comments in one row/cell. 
# If that is the case, you need to be careful not to count it twice/
# overwrite the reason/decide which is the primary cause for exclusion.

# Add flag
excl_conc <- excl_conc %>% 
  mutate(REASON = ifelse(str_detect(COMMENT, "Pre first dose sample"), 
                         "Pre first dose sample", NA),
         REASON = ifelse(str_detect(COMMENT, "Randomly selected for illustration"), 
                         "Randomly selected for illustration", REASON)
         )

# Total n concentration 
total_conc <- rawdata %>% filter(EVID==0)

# Generate summary and tweak to presentable table
summaryExclConc <- 
  excl_conc %>%
  group_by(REASON, BLQ) %>% 
  summarize(n = n(), 
            `(%)` = round(100 * (n / nrow(total_conc)), digits=1)) %>% 
  ungroup() %>% 
  mutate(BLQ = factor(BLQ, levels = c(0,1), labels = c("Non-BLQ", "BLQ")))
  
summaryExclConc
```


```{r, include=F}
# Write summary table to file
if(params$print_results){
  write.csv(summaryExclConc,
            file = file.path(directories[["res_eda_dir"]], "omittedConcDataSummary.csv"),
            row.names = F)
}

# Full listing to file
listingExcluded <- 
  excl_conc %>%  # in this case using excl_conc as no doses were excluded
  # excluded %>% 
  select(NMSEQSID,TIME,TAPD,DV,AMT,EVID,MDV,BLQ,DOSE,REASON)

if(params$print_results){
  write.csv(listingExcluded,
            file = file.path(directories[["res_eda_dir"]], "omittedDataListing.csv"),
            row.names = F)
}
```


### How much of the non-BLQ excluded data are from sparse samples:
```{r}
# Note, in this example the sparse data is those without a OCC assignment
summarySparse <- 
  excl_conc %>%
  filter(is.na(OCC) & BLQ==0) %>% 
  group_by(REASON) %>% 
  summarize(n = n(), 
            `(%)` = round(100 * (n / nrow(total_conc) ), digits=1))
summarySparse
```



# Review of comments not leading to exclusion
```{r}
# Use the dataset 'data' for this since all C=="C" is commented out
unique(data$COMMENT)

data %>%
  filter(! (COMMENT %in% c(" ", "") | is.na(COMMENT))) %>% 
  group_by(EVID, COMMENT) %>% 
  summarize(n = n(), 
            percentOfTotal = round(100 * (n / nrow(data)), digits=2))
```

